---
title: "Metrics II - Homework 8"
author: "Farnam Taheri"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: readable
    highlight: kate
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
    code_folding: hide
    css: styles.css
    includes:
      before_body: navbar.html
editor_options:
  markdown:
    wrap: 72
---

```{r setup, warning=FALSE, message=FALSE, results=FALSE}
# Setup chunk to load required packages and set global options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)

# Install necessary packages if not already installed
if (!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")
if (!requireNamespace("synthdid", quietly = TRUE)) devtools::install_github("synth-inference/synthdid")
if (!requireNamespace("plotly", quietly = TRUE)) install.packages("plotly")
if (!requireNamespace("DT", quietly = TRUE)) install.packages("DT")
if (!requireNamespace("htmltools", quietly = TRUE)) install.packages("htmltools")
if (!requireNamespace("stargazer", quietly = TRUE)) install.packages("stargazer")

# Load the packages
library(synthdid)
library(plotly)
library(DT)
library(htmltools)
library(stargazer)
```

# [Part I: Synthetic Control]{.smallcaps} {#part1}

**In this exercise we implement the synthetic DID methods approach.
Synthetic DID combines both the concepts of DID and synthetic control.
In Stata, you should use the sdid package
(<https://github.com/Daniel-Pailanir/sdid>) and in R use the synthdid
package (<https://synth-inference.github.io/synthdid/>).**

## [Preliminaries]{.smallcaps}

### Define the Required Functions

We will define functions for estimating synthetic control, computing
standard errors, visualizing results, and generating interactive tables.

#### Prepare Synthetic Data

```{r}
# Function to Prepare Data for Synthetic Control Analysis

# This function takes a dataset and the names of the columns that represent units, time, outcomes, and treatment.
# It formats the data into a structure suitable for synthetic control analysis using the synthdid package.
prepare_synthetic_data <- function(data, unit_col, time_col, outcome_col, treatment_col) {
  # Inform the user that data preparation is starting
  message("Preparing data for synthetic control analysis...")
  
  # Attempt to prepare the data and handle any errors that may occur
  tryCatch({
    # Convert the data into the format required by the synthdid package
    setup <- panel.matrices(data, unit = unit_col, time = time_col, outcome = outcome_col, treatment = treatment_col)
    # Return the prepared data
    return(setup)
  }, error = function(e) {
    # Provide a more informative error message if preparation fails
    stop("An error occurred while preparing the data: ", e$message)
  })
}
```

#### Estimate Synthetic Control

```{r}
# Function to Estimate Synthetic Control with HTML Output

# This function takes the prepared data and uses the synthdid_estimate function to estimate the synthetic control.
# It provides an HTML output for interpretation of the results.
estimate_synthetic_control <- function(setup, font_family = "Arial", font_size = "16px", color = "black") {
  # Inform the user that the synthetic control estimation is starting
  message("Estimating synthetic control...")
  
  # Attempt to estimate the synthetic control and handle any errors that may occur
  tryCatch({
    # Use the sc_estimate function from the synthdid package to estimate the synthetic control
    synth_control <- sc_estimate(setup$Y, setup$N0, setup$T0)
    
    # Generate HTML output with inline CSS for styling
    # The output includes a title and a paragraph with the point estimate
    html_output <- paste0(
      "<h2 style='font-family: ", font_family, "; color: ", color, "; font-size: 24px;'>Synthetic Control Estimation Results</h2>",
      "<p style='font-family: ", font_family, "; color: ", color, "; font-size: ", font_size, ";'><strong>Point Estimate:</strong> ", 
      round(synth_control, 3), "</p>",
      "<p style='font-family: ", font_family, "; color: gray; font-size: 18px;'>This point estimate represents the counterfactual outcome that would have been observed in the absence of the treatment. It is calculated as a weighted average of the control units.</p>"
    )
    
    # Print the HTML output to the console
    cat(html_output)
    
    # Return the estimated synthetic control object
    return(synth_control)
  }, error = function(e) {
    # Provide a more informative error message if estimation fails
    stop("An error occurred during the estimation: ", e$message)
  })
}
```

#### Estimate Synthetic Diff-in-Diff

```{r}
# Function to Estimate SynthDID with HTML-formatted Interpretations and Styling Options
estimate_synthdid <- function(setup, 
                              font_family = "Arial", 
                              font_size = "16px", 
                              font_color = "black") {
  # Inform the user that the SynthDID estimation is starting
  message("Estimating SynthDID...")
  
  # Attempt to estimate the SynthDID and handle any errors that may occur
  tryCatch({
    # Use the synthdid_estimate function to estimate the synthetic control
    synthdid_result <- synthdid_estimate(setup$Y, setup$N0, setup$T0)
    
    # Extract the estimate and weights from the result
    estimate <- synthdid_result[1]
    weights <- attr(synthdid_result, "weights")
    
    # Calculate effective sample sizes
    effective_N0 <- sum(weights$omega)
    effective_T0 <- sum(weights$lambda)
    
    # Create HTML-formatted interpretations based on the estimate
    interpretation <- if (estimate < 0) {
      "<p>The negative SynthDID estimate suggests a decrease in the outcome variable due to the treatment.</p>"
    } else if (estimate > 0) {
      "<p>The positive SynthDID estimate suggests an increase in the outcome variable due to the treatment.</p>"
    } else {
      "<p>The SynthDID estimate is zero, indicating no effect of the treatment on the outcome variable.</p>"
    }
    
    # Add details about effective sample sizes to the interpretation
    interpretation <- paste0(
      "<div style='font-family:", font_family, "; font-size:", font_size, "; color:", font_color, ";'>",
      interpretation,
      "<p>Effective N0/N0 ratio: <strong>", round(effective_N0 / setup$N0, 2), "</strong> ",
      "suggests that <strong>", effective_N0, "</strong> out of <strong>", setup$N0, "</strong> ",
      "control units are effectively used in constructing the synthetic control.</p>",
      "<p>Effective T0/T0 ratio: <strong>", round(effective_T0 / setup$T0, 2), "</strong> ",
      "indicates that <strong>", effective_T0, "</strong> out of <strong>", setup$T0, "</strong> ",
      "pre-treatment periods are effectively used.</p>",
      "<p>N1 and T1 represent the number of treated units and post-treatment periods, respectively.</p>",
      "</div>"
    )
    
    # Print the HTML-formatted interpretation
    cat(interpretation, '\n')
    
    # Return the estimated SynthDID object
    return(synthdid_result)
  }, error = function(e) {
    # Provide a more informative error message if estimation fails
    stop("An error occurred during the SynthDID estimation: ", e$message)
  })
}

```

#### Plot Synthetic DiD Control

```{r}
# Function to Plot Synthetic Control 

# This function takes the estimated synthetic control object and an optional title to create an interactive plot.
plot_synthdid_control <- function(synth_control, title = "Synthetic Control Results", legend_position = "bottom", font_family = "serif", font_size = 12, font_color = "black") {
  tryCatch({
    p <- plot(synth_control) +
      ggtitle(title) +
      theme_minimal() +
      theme(text = element_text(family = font_family, size = font_size, color = font_color), # Set font for the entire plot
            plot.title = element_text(hjust = 0.5, face = "bold"),
            legend_position = legend_position,
            legend_title = element_blank(), # Remove the legend title
            legend_background = element_rect(fill = "white", colour = "black"), # Customize legend background
            legend_key = element_rect(fill = "white", colour = "black")) # Customize legend keys
    
    ggplotly(p)
  }, error = function(e) {
    stop("An error occurred while plotting: ", e$message)
  })
}

```

#### Summary HTML

```{r}
summary_html <- function(estimate_syhtetics, caption = "Summary of Synthetic Estimate", font_family = "Arial", font_size = "16px", color = "black") {
  
  # Extract the relevant parts from the summary output
  estimate <- summary(estimate_syhtetics)$estimate
  controls <- summary(estimate_syhtetics)$controls
  periods <- summary(estimate_syhtetics)$periods
  dimensions <- summary(estimate_syhtetics)$dimensions
  
  # Start HTML content with enhanced styling
  html_content <- paste0("<div style='font-family: ", font_family, "; font-size: ", font_size, "; color: ", color, "; background-color: #f8f9fa; padding: 10px;'>")
  
  # Add caption
  html_content <- paste0(html_content, "<h4 style='text-align: center;'>", caption, "</h4>")
  
  # Add estimate with explanation
  html_content <- paste0(html_content, "<p style='text-align: center;'><strong>Estimated Treatment Effect:</strong> ", round(estimate, 2), 
                         ". This value represents the estimated impact of the intervention.</p>")
  
  # Add controls with enhanced visualization
  html_content <- paste0(html_content, "<h5 style='color: #17a2b8; text-align: center;'>Control Unit Weights:</h5>",
                         "<p style='text-align: center;'>The weights of the control units indicate their relative importance in constructing the synthetic control group. Higher weights suggest a greater similarity to the treated unit.</p>",
                         "<table style='width: 100%; border-collapse: collapse; margin-left: auto; margin-right: auto;'>",
                         "<tr style='background-color: #e9ecef;'><th style='text-align: center;'>Control Unit</th><th style='text-align: center;'>Weight</th></tr>")
  for (control in rownames(controls)) {
    html_content <- paste0(html_content, "<tr><td style='text-align: center;'>", control, "</td><td style='text-align: center;'>", round(controls[control, ], 3), "</td></tr>")
  }
  html_content <- paste0(html_content, "</table>")
  
  # Add periods with enhanced visualization
  html_content <- paste0(html_content, "<h5 style='color: #17a2b8; text-align: center;'>Period Weights:</h5>",
                         "<p style='text-align: center;'>The period weights reflect the importance of each time period in the pre-intervention phase for constructing the synthetic control.</p>",
                         "<ul style='list-style-type: none; padding: 0; text-align: center;'>")
  for (period in names(periods)) {
    html_content <- paste0(html_content, "<li>", period, ": ", round(periods[period], 3), "</li>")
  }
  html_content <- paste0(html_content, "</ul>")
  
  # Add dimensions with explanation
  html_content <- paste0(html_content, "<h5 style='color: #17a2b8; text-align: center;'>Dimensions:</h5>",
                         "<p style='text-align: center;'>These numbers provide information about the structure of the dataset used for the analysis:</p>",
                         "<ul style='list-style-type: none; padding: 0; text-align: center;'>")
  for (dimension in names(dimensions)) {
    html_content <- paste0(html_content, "<li><strong>", dimension, ":</strong> ", dimensions[dimension], "</li>")
  }
  html_content <- paste0(html_content, "</ul>")
  
  # End HTML content
  html_content <- paste0(html_content, "</div>")
  
  # Print the HTML content
  cat(html_content)
}

```

#### Perform Placebo Test

```{r}
# Function to Perform a Placebo Testپ
perform_placebo_test <- function(setup, years_shift, 
                                 font_family = "serif", 
                                 font_size = "16px", 
                                 font_color = "black") {
  # Inform the user that the placebo test is starting
  message("Performing placebo test...")
  
  # Attempt to perform the placebo test and handle any errors that may occur
  tryCatch({
    # Create a copy of the setup object and shift the treatment period for the placebo test
    placebo_setup <- setup
    placebo_setup$T0 <- placebo_setup$T0 - years_shift
    
    # Estimate the placebo effect using the synthdid_estimate function
    placebo_estimate <- synthdid_estimate(placebo_setup$Y, placebo_setup$N0, placebo_setup$T0)
    
    # Extract the estimate and weights from the result
    estimate <- placebo_estimate[1]
    weights <- attr(placebo_estimate, "weights")
    
    # Calculate effective sample sizes
    effective_N0 <- sum(weights$omega)
    effective_T0 <- sum(weights$lambda)
    
    # Create an enhanced HTML-formatted interpretation based on the estimate
    interpretation <- if (estimate < 0) {
      "<p>The observed negative placebo estimate suggests a potential decrease in the outcome variable, which could imply an overestimation of the treatment effect when using the actual treatment year.</p>"
    } else if (estimate > 0) {
      "<p>The observed positive placebo estimate indicates a potential increase in the outcome variable, suggesting an underestimation of the treatment effect if this were the actual treatment year.</p>"
    } else {
      "<p>The placebo estimate is zero, indicating no discernible difference between the synthetic and actual outcomes, thus supporting the robustness of the original treatment effect estimation.</p>"
    }
    
    # Add details about effective sample sizes to the interpretation
    interpretation <- paste0(
      "<div style='font-family:", font_family, "; font-size:", font_size, "; color:", font_color, ";'>",
      interpretation,
      "<p>The Effective N0/N0 ratio of <strong>", round(effective_N0 / setup$N0, 2), "</strong> ",
      "reflects the proportion of control units that significantly contribute to constructing the synthetic control, with an effective count of <strong>", round(effective_N0, 3), "</strong> out of <strong>", setup$N0, "</strong> total units.</p>",
      "<p>Similarly, the Effective T0/T0 ratio of <strong>", round(effective_T0 / setup$T0, 2), "</strong> ",
      "reveals the extent to which pre-treatment periods are utilized, with <strong>", effective_T0, "</strong> out of <strong>", setup$T0, "</strong> periods being effectively used in the analysis.</p>",
      "<p>These ratios provide insight into the representativeness and reliability of the synthetic control in relation to the actual data.</p>",
      "</div>"
    )
    
    # Print the HTML-formatted interpretation
    cat(interpretation, '\n')
    
    # Return the estimated placebo effect object
    return(placebo_estimate)
  }, error = function(e) {
    # Provide a more informative error message if the placebo test fails
    stop("An error occurred during the placebo test: ", e$message)
  })
}

```

#### Analyze Weights

```{r}
generate_interactive_table <- function(
    data, top_caption = NULL, col_names = names(data),
    options = list(pageLength = 5, scrollX = TRUE),
    row_colors = NULL, round_digits = 2, background_color = "white",
    bottom_footnote = NULL, font_family = "Arial", font_size = "16px", color = "black") {
  
  # Ensure the input data is a data frame to prevent errors
  stopifnot(is.data.frame(data))
  
  # Round the numeric columns in the data frame if round_digits is specified
  if (!is.null(round_digits)) {
    numeric_columns <- sapply(data, is.numeric)
    data[numeric_columns] <- lapply(data[numeric_columns], round, round_digits)
  }
  
  # Set custom column names if provided, otherwise use the default column names
  if (!is.null(col_names)) {
    names(data) <- col_names
  }
  
  # Create a datatable object with the provided data and additional formatting options
  table <- DT::datatable(
    data, 
    caption = htmltools::tags$caption(
      style = paste('caption-side: top; text-align: center; font-family:', font_family, '; font-size:', font_size, '; color:', color, ';'),
      top_caption),
    options = options
  ) %>% DT::formatStyle(columns = names(data), backgroundColor = background_color, textAlign = 'center')
  
  # Apply row colors if provided, enhancing the visual distinction between rows
  if (!is.null(row_colors)) {
    table <- table %>% DT::formatStyle(0, target = 'row', backgroundColor = row_colors)
  }
  
  # Add a footnote at the bottom of the table if provided
  if (!is.null(bottom_footnote)) {
    table <- table %>%
      DT::addFooter(
        htmltools::tags$div(
          style = paste('text-align: center; margin-top: 10px; font-family:', font_family, '; font-size:', font_size, '; color:', color, ';'),
          htmltools::HTML(bottom_footnote)
        )
      )
  }
  
  # Return the interactive table object for rendering in the UI
  return(table)
}

```

### Load Data

Load the dataset `california_prop99` provided by the `synthdid` package
and prepare it for analysis.

```{r}
# Load the dataset
data('california_prop99', package = 'synthdid')
```

```{r message=FALSE}
# Prepare the data
setup <- prepare_synthetic_data(
  california_prop99,
  unit_col = "State",
  time_col = "Year",
  outcome_col = "PacksPerCapita",
  treatment_col ="treated")
```

```{r}
generate_interactive_table(
  data = california_prop99, 
  top_caption = "California Proposition 99 Data", 
  row_colors = 'skyblue', 
  round_digits = 3, 
  background_color = 'aliceblue',
  font_family = "Serif"
)
```

## [1. Differences Between SCM, DID, and SynthDID Estimation.]{.smallcaps}

**First explain the differences between synthetic control, DID, and
synthetic DID estimation.**

**Difference-in-Differences (DID):**

-   **Methodology**: DID estimates the effect of a treatment by
    comparing the changes in outcomes over time between a treated group
    and a control group.

-   **Assumption**: The main assumption is that in the absence of
    treatment, the difference between the treated and control groups
    would have remained constant over time (parallel trends assumption).

-   **Use Case**: Suitable when there is a clear treatment and control
    group, and the parallel trends assumption holds.

**Synthetic Control Method (SCM):**

-   **Methodology**: SCM constructs a synthetic version of the treated
    unit by weighting control units to best match the characteristics of
    the treated unit before the treatment. The difference in outcomes
    between the treated unit and its synthetic counterpart after
    treatment provides an estimate of the treatment effect.

-   **Assumption**: The synthetic control group should closely mimic the
    treated unit’s characteristics before the intervention, thus serving
    as a better counterfactual than a single control unit.

-   **Use Case**: Effective in cases with a single treated unit and
    multiple control units, especially when the parallel trends
    assumption is questionable.

**Synthetic DID (SynthDID):**

-   **Methodology**: SynthDID combines DID and SCM to leverage both
    temporal and cross-sectional variations, offering a robust estimator
    that can relax the parallel trends assumption by creating a more
    accurate counterfactual using synthetic controls.

-   **Assumption**: SynthDID uses the strengths of both methods to
    handle situations where the parallel trends assumption might be
    relaxed.

-   **Use Case**: Suitable for settings where both DID and SCM have
    advantages, providing a more reliable estimate when the parallel
    trends assumption is weak.

## [2. Estimate and Visualize the Synthetic Control]{.smallcaps}

**We will replicate the main synthetic control effect of California’s
Prop 99 on cigarette consumption (you can see [ADH
2010](https://web.stanford.edu/~jhain/Paper/JASA2010.pdf "ADH 2010") for
your reference). Setup the data following the vignette for the package.
Report the treatment effect for the synthetic control. (note, this is
not the main synthdid estimator).**

```{r message=FALSE, results='asis'}
synth_control_estimate <- estimate_synthetic_control(setup, font_size = "18px", font_family = 'serif')
```

```{r}
synthdid::synthdid_plot(synth_control_estimate, line.width = 1)
```

## [3. Estimating the Effect Using the SynthDID Estimator]{.smallcaps}

**Now, estimate the effect using the synthdid estimator. Report the
treatment effect and compare to the synthetic control effect.**

In this part, we estimate the effect using the SynthDID estimator and
compare it to the synthetic control results

```{r message=FALSE, results='asis'}
synth_did_estimate <- estimate_synthdid(setup, font_family='serif', font_size = "18px")
```

```{r}
plot_synthdid_control(
  synth_did_estimate,
  "Synthetic Diff-in-Diff for California's Prop 99 on Cigarette Sales",
  )
```

## [4. Analyzing Weights for the Synthetic Control]{.smallcaps}

**Take a look at the weights for the synthetic control. What variables
are most important in the synthetic control? How do these weights
compare to the weights in the synthetic did estimator? (In R, this is
done by looking at the summary command for the estimate)**

```{r , results='asis'}
summary_html(synth_control_estimate, font_family='serif', caption = "Summary of Synthetic Control Method", font_size = "18px")
```

```{r}
synthdid_units_plot(synth_control_estimate)
```

```{r message=FALSE, results='asis'}
summary_html(synth_did_estimate, font_family='serif', caption = "Summary of Synthetic DiD Method", font_size = "18px")
```

```{r}
synthdid_units_plot(synth_did_estimate)
```

# [Part II: DID Analysis]{.smallcaps} {#part2}

**In this exercise, we want to estimate the effect of the 1398 Southwest
flood using DID. Use Household Expenditure and Income Surveys and answer
the following questions:**

## [Preliminaries]{.smallcaps}

```{r, results=FALSE}
# Load the required libraries
library(DBI)
library(RODBC)
library(foreign) 
library(dplyr)
library(data.table)
library(ggplot2)
library(tidyverse)
library(readxl)
```

### [Define Functions]{.smallcaps}

#### Prepare and Save Data

```{r}
# Function to prepare and save urban and rural datasets
prepare_and_save_data <- function(name, urban_df, rural_df) {
  # Add 'urban' column
  urban_df$urban <- 1
  rural_df$urban <- 0
  
  # Combine Urban and Rural data.frames
  combined_df <- rbind(urban_df, rural_df)
  
  # Save the combined dataset in the ./export folder
  saveRDS(combined_df, file = paste0("./export/", name, ".RDS"))
}

# List of data.frame names
data_frames <- c("97P1", "97P3S01", "97P3S02", "97Data", "97P3S03", "97P3S04", "97P3S05",
                 "97P3S06", "97P3S07", "97P3S08", "97P3S09", "97P3S11", "97P3S12",
                 "97P4S01", "97P4S02", "97P4S03", "97P4S04",
"98P1", "98P3S01", "98P3S02", "98Data", "98P3S03", "98P3S04", "98P3S05",
                 "98P3S06", "98P3S07", "98P3S08", "98P3S09", "98P3S11", "98P3S12",
                 "98P4S01", "98P4S02", "98P4S03", "98P4S04"
)

# Check if the data.frame is empty
is_empty <- function(df) {
  return(nrow(df) == 0)
}
```

#### Load Processed Data

```{r}
load_processed_data <- function(file_path) {
  if (!file.exists(file_path)) {
    cat("The file does not exist at the specified path: ", file_path, "\n")
    return(NULL)
  }
  tryCatch({
    data <- readRDS(file_path)
    cat("File loaded successfully.\n")
    return(data)
  }, error = function(e) {
    cat("Error loading the file: ", e$message, "\n")
    return(NULL)
  })
}
```

#### Filter Treatment and Control

```{r}
filter_treatment_control <- function(data) {
  # Define treatment and control district codes
  lorestan_treatment_districts <- c("02", "10", "04", "05", "03", "06", "08", "09")
  khuzestan_treatment_districts <- c("03", "09", "08", "11", "14", "12")
  
  lorestan_control_districts <- c("01", "07", "15")
  khuzestan_control_districts <- c("01", "21", "02", "06", "20", "18", "04", "07", "17", "05", "13", "10", "19")
  
  # Define province codes
  lorestan_code <- "15"
  khuzestan_code <- "06"
  
  # Filter data for treatment groups
  treatment_data <- data %>%
    filter((substr(Address, 2, 3) == khuzestan_code & substr(Address, 4, 5) %in% khuzestan_treatment_districts) |
           (substr(Address, 2, 3) == lorestan_code & substr(Address, 4, 5) %in% lorestan_treatment_districts))
  
  # Filter data for control groups
  control_data <- data %>%
    filter((substr(Address, 2, 3) == khuzestan_code & substr(Address, 4, 5) %in% khuzestan_control_districts) |
           (substr(Address, 2, 3) == lorestan_code & substr(Address, 4, 5) %in% lorestan_control_districts))
  
  return(list(treatment = treatment_data, control = control_data))
}

```

#### Apply Custom Styling

```{r, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

# Function to apply custom styling to kable tables
apply_custom_styling <- function(data_frame) {
  kable(data_frame, format = "html", escape = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                  full_width = F, 
                  position = "center", # Updated to center the table
                  font_size = 16,
                  fixed_thead = T) %>%
    row_spec(0, bold = T, color = "midnightblue", background = "aliceblue") 
}
```

#### Calculate Income

```{r}
# Function to categorize, rename columns, calculate incomes, and generate totals
calculate_income_by_category <- function(data_list, dataset_names) {
  # Define the patterns for different categories with meaningful names
  categories <- list(
    WageAndSalaryIncome = c("P4S01$"),  # Income from wage and salary jobs
    SelfEmploymentIncome = c("P4S02$"),  # Income from self-employment
    HouseholdIncome = c("P4S03$"),  # Miscellaneous household incomes
    TransfersAndOtherIncome = c("P4S04$")   # Transfers and other incomes
  )
  
  # Initialize results dataframe
  results <- data.frame(Category = character(), Dataset = character(), TotalIncome = numeric(), PerCapitaIncome = numeric(), stringsAsFactors = FALSE)
  
  # Initialize cumulative income and population counts for each category
  category_totals <- data.frame(Category = names(categories), TotalIncome = rep(0, length(categories)), TotalPopulation = rep(0, length(categories)), stringsAsFactors = FALSE)
  
  # Loop through each dataset
  for (i in 1:length(data_list)) {
    df <- data_list[[i]]
    dataset_name <- dataset_names[[i]]
    
    # Determine the category of the current dataset
    category <- NA
    for (cat in names(categories)) {
      if (any(sapply(categories[[cat]], function(pattern) grepl(pattern, dataset_name)))) {
        category <- cat
        break
      }
    }
    
    # Skip datasets that do not match any category
    if (is.na(category)) next
    
    # Rename columns based on the dataset name pattern
    if (grepl("P4S1$", dataset_name)) {
      names(df) <- c("Address", "MemberID", "IsEmployed", "JobCode", "MainWorkplaceCode", "SectorCode", "DailyWorkHours", "WeeklyWorkDays", "MonthlyGrossIncome", "AnnualGrossIncome", "MonthlyWages", "AnnualWages", "MonthlyNonWageBenefits", "AnnualNonWageBenefits", "MonthlyNetIncome", "AnnualNetIncome")
      df$Value <- df$AnnualNetIncome
    } else if (grepl("P4S2$", dataset_name)) {
      names(df) <- c("Address", "MemberID", "IsEmployed", "JobCode", "MainWorkplaceCode", "EmployerType", "AgricultureType", "DailyWorkHours", "WeeklyWorkDays", "MonthlyIncome", "SeedWaterFertilizerCost", "ToolCost", "JobCost", "TaxCost", "GrossReceipts", "NetIncome")
      df$Value <- df$NetIncome
    } else if (grepl("P4S3$", dataset_name)) {
      names(df) <- c("Address", "MemberID", "PensionIncome", "RentIncome", "SavingsIncome", "ScholarshipIncome", "SaleIncome", "TransferIncome")
      df$Value <- rowSums(df[, c("PensionIncome", "RentIncome", "SavingsIncome", "ScholarshipIncome", "SaleIncome", "TransferIncome")], na.rm = TRUE)
    } else if (grepl("P4S4$", dataset_name)) {
      names(df) <- c("Address", "MemberID", "SubsidyRecipients", "SubsidyMonths", "TotalSubsidy")
      df$Value <- df$TotalSubsidy
    }
    
    # Calculate the total income and population for the current dataset
    total_income <- sum(df$Value, na.rm = TRUE)
    total_population <- nrow(df)
    per_capita_income <- ifelse(total_population > 0, total_income / total_population, NA)
    
    # Append the results to the results dataframe
    results <- rbind(results, data.frame(Category = category, Dataset = dataset_name, TotalIncome = total_income, PerCapitaIncome = per_capita_income, stringsAsFactors = FALSE))
    
    # Update the cumulative totals for the category
    category_totals[category_totals$Category == 'WageAndSalaryIncome', "TotalIncome"] <- category_totals[category_totals$Category == 'WageAndSalaryIncome', "TotalIncome"] + total_income
    category_totals[category_totals$Category == 'SelfEmploymentIncome', "TotalPopulation"] <- category_totals[category_totals$Category == 'SelfEmploymentIncome', "TotalPopulation"] + total_population
  }
  
  # Calculate per capita income for each category
  category_totals$PerCapitaIncome <- with(category_totals, ifelse(TotalPopulation > 0, TotalIncome / TotalPopulation, NA))
  
  # Calculate the sum of incomes for all categories
  sum_total_income <- sum(category_totals$TotalIncome)
  sum_per_capita_income <- sum(category_totals$PerCapitaIncome, na.rm=TRUE)
  
  # Combine the sums into a new data frame
  sums <- data.frame(
    SumTotalIncome = sum_total_income,
    SumPerCapitaIncome = sum_per_capita_income
  )
  
  # Add the sums to the return list
  return(list(
    results = results,
    category_totals = category_totals,
    sums = sums
  ))
}

```

#### Calculate Expenditure

```{r}
# Function to categorize, rename columns, calculate expenditures, and generate totals
calculate_exp_by_category <- function(data_list, dataset_names) {
  # Define the patterns for different categories
  categories <- list(
    category1 = c("P3S01$", "P3S02$"),
    category2 = c("P3S03$", "P3S04$"),
    category3 = c("P3S0[3,5-9,11,12]$", "P3S13$", "P3S14$")
  )
  
  # Initialize results dataframe
  results <- data.frame(Category = character(), Dataset = character(), TotalExpenditure = numeric(), PerCapitaExpenditure = numeric(), stringsAsFactors = FALSE)
  
  # Initialize cumulative expenditure and population counts for each category
  category_totals <- data.frame(Category = names(categories), TotalExpenditure = rep(0, length(categories)), TotalPopulation = rep(0, length(categories)), stringsAsFactors = FALSE)
  
  # Loop through each dataset
  for (i in 1:length(data_list)) {
    df <- data_list[[i]]
    dataset_name <- dataset_names[[i]]
    
    # Determine the category of the current dataset
    category <- NA
    for (cat in names(categories)) {
      if (any(sapply(categories[[cat]], function(pattern) grepl(pattern, dataset_name)))) {
        category <- cat
        break
      }
    }
    
    # Skip datasets that do not match any category
    if (is.na(category)) next
    
    # Rename columns based on the dataset name pattern
    if (grepl("P3S0[1-2]$", dataset_name)) {
      names(df) <- c("Address", "ItemCode", "PurchaseMethod", "WeightGram", "WeightKilo", "UnitPrice", "Value")
      df$Value <- ifelse(is.na(df$Value), df$WeightKilo * df$UnitPrice, df$Value)
    } else if (grepl("P3S04$", dataset_name)) {
      names(df) <- c("Address", "ItemCode", "DepositAmount", "PurchaseMethod", "Value")
      df$Value <- ifelse(is.na(df$Value), df$DepositAmount, df$Value)
    } else if (grepl("P3S13$", dataset_name)) {
      names(df) <- c("Address", "ItemCode", "InsuredPeopleCountLoanAmount", "LoanLocation", "PurchaseMethod", "PurchaseCost", "SecondHandSale")
      df$Value <- ifelse(is.na(df$Value), df$PurchaseCost, df$Value)
    } else if (grepl("P3S14$", dataset_name)) {
      names(df) <- c("Address", "ItemCode", "PurchaseMethod", "PurchaseCost", "SecondHandSale")
      df$Value <- ifelse(is.na(df$Value), df$PurchaseCost, df$Value)
    } else if (grepl("P3S0[3,5-9,11,12]$", dataset_name)) {
      names(df) <- c("Address", "ItemCode", "PurchaseMethod", "Value")
    }
    
    # Calculate the total expenditure and population for the current dataset
    total_expenditure <- sum(df$Value, na.rm = TRUE)
    total_population <- nrow(df)
    per_capita_expenditure <- ifelse(total_population > 0, total_expenditure / total_population, NA)
    
    # Append the results to the results dataframe
    results <- rbind(results, data.frame(Category = category, Dataset = dataset_name, TotalExpenditure = total_expenditure, PerCapitaExpenditure = per_capita_expenditure, stringsAsFactors = FALSE))
    
    # Update the cumulative totals for the category
    category_totals[category_totals$Category == category, "TotalExpenditure"] <- category_totals[category_totals$Category == category, "TotalExpenditure"] + total_expenditure
    category_totals[category_totals$Category == category, "TotalPopulation"] <- category_totals[category_totals$Category == category, "TotalPopulation"] + total_population
  }
  
  # Calculate per capita expenditure for each category
  category_totals$PerCapitaExpenditure <- with(category_totals, ifelse(TotalPopulation > 0, TotalExpenditure / TotalPopulation, NA))
  
    
  # Calculate the sum of expenditures for all categories
  sum_total_expenditure <- sum(category_totals$TotalExpenditure)
  sum_per_capita_expenditure <- sum(category_totals$PerCapitaExpenditure,        na.rm=TRUE)
  
  # Combine the sums into a new data frame
  sums <- data.frame(
    SumTotalExpenditure = sum_total_expenditure,
    SumPerCapitaExpenditure = sum_per_capita_expenditure
  )
  
  # Add the sums to the return list
  return(list(
    results = results,
    category_totals = category_totals,
    sums = sums
  ))
}
```

### Data Cleaning

#### Transform the `MDB` file.

```{r message=FALSE, warning=FALSE, results=FALSE}
# Define the base file path
base_path <- "./Data/"

# Loop over the years 97 and 98
for (year in c("Data97", "Data98")) {
  # Construct the full file path for the current year
  db <- file.path(base_path, year)
  
  # Connect to the Access database using odbcDriverConnect
  conn_str <- paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)};DBQ=", db)
  channel <- odbcDriverConnect(conn_str)
  
  # Get the names of all tables in the database
  names <- subset(sqlTables(channel), TABLE_TYPE == "TABLE", TABLE_NAME)[[1]]
  
  # Iterate over each table name
  for (i in names) {
    print(i)
    nam <- paste(i)
    
    # Fetch the table and assign it to a variable with the same name as the table
    assign(nam, sqlFetch(channel, i))
  }
  
  # Clean up the environment
  rm("channel", "i", "nam", "names", "db")
  
  # Save the workspace image for the current year
  save.image(file=paste0("./HEIS/", substr(year, 5, 6), ".RData"))
}
```

```{r}
load("./HEIS/97.RData")
```

```{r}
# Apply the check before combining the data.frames
for (name in data_frames) {
  urban_df <- get(paste0("U", name))
  rural_df <- get(paste0("R", name))
  
  if (is_empty(urban_df) || is_empty(rural_df)) {
    cat("Data.frame ", name, " is empty and cannot be processed.\n")
  } else {
    prepare_and_save_data(name, urban_df, rural_df)
  }
}
```

#### Load the Data

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Function to load processed data
load_processed_data <- function(file_path) {
  readRDS(file_path)
}

# List of dataset suffixes
dataset_suffixes <- c("97P1", "97P3S01", "97P3S02", "97Data", "97P3S03", "97P3S04", "97P3S05",
                      "97P3S06", "97P3S07", "97P3S08", "97P3S09", "97P3S11", "97P3S12",
                      "97P4S01", "97P4S02", "97P4S03", "97P4S04")

# Initialize lists to store dataframes and their names
data_list <- list()
dataset_names <- list()

# Loop through each suffix and load the corresponding RDS file
for (suffix in dataset_suffixes) {
  file_path <- paste0("./export/Data", suffix, ".RDS")
  dataset_name <- paste0("data_", suffix)
  data_list[[suffix]] <- load_processed_data(file_path)
  dataset_names[[suffix]] <- dataset_name
}

# List of dataset suffixes
dataset_suffixes_98 <- c("97P1", "97P3S01", "97P3S02", "97Data", "97P3S03", "97P3S04", "97P3S05",
                      "97P3S06", "97P3S07", "97P3S08", "97P3S09", "97P3S11", "97P3S12",
                      "97P4S01", "97P4S02", "97P4S03", "97P4S04")

# Initialize lists to store dataframes and their names
data_list <- list()
dataset_names <- list()

# Loop through each suffix and load the corresponding RDS file
for (suffix in dataset_suffixes) {
  file_path <- paste0("./export/Data", suffix, ".RDS")
  dataset_name <- paste0("data_", suffix)
  data_list[[suffix]] <- load_processed_data(file_path)
  dataset_names[[suffix]] <- dataset_name
}
```

```{r}
# Initialize lists to store filtered datasets
filtered_treatment_list <- list()
filtered_control_list <- list()

# Loop through each dataset and apply the filtering function
for (suffix in dataset_suffixes) {
  # Apply the filter function to each dataset
  filtered_data <- filter_treatment_control(data_list[[suffix]])
  
  # Store the filtered data into respective lists
  filtered_treatment_list[[suffix]] <- filtered_data$treatment
  filtered_control_list[[suffix]] <- filtered_data$control
}
# Now, filtered_treatment_list and filtered_control_list contain the treatment and control data for each dataset.
```

## [1. Exogenous Shock]{.smallcaps}

**Do you find the Lorestan & Khuzestan flood as an exogenous shock?
explain.**

The Lorestan & Khuzestan flood of 1398 can be seen as an exogenous shock
for a few key reasons. Firstly, the flood was unexpected and caused by
unusual weather patterns that were beyond the control or prediction of
the local people and authorities. Such extreme weather events are rare
and difficult to foresee, which makes the flood an external and sudden
event.

Secondly, the flood's occurrence was not linked to the characteristics
or actions of the households in Lorestan & Khuzestan. The regions did
not have any specific features that would make them more prone to
flooding compared to other areas. This means that the flood's impact was
not influenced by factors like the residents' income levels, spending
habits, or types of housing, making it a random event that affected
everyone equally, regardless of their situation.

Lastly, the flood's clear location and timing help to classify it as an
exogenous shock. The flood affected specific areas at a specific time,
which makes it easy to identify who was impacted and who was not. This
clear distinction allows for a more accurate analysis of how the flood
affected household expenditure and income, supporting the idea that the
flood was an external shock that disrupted normal life in an unforeseen
way.

## [2. Data Summary]{.smallcaps}

**Import the data, describe and summarize the variables of P4S2 tables
for urban and rural residents at 1397 & 1398.**

```{r, results='asis'}
generate_interactive_table(data = R97P4S02,
                           top_caption="R97P4S02",
                           font_family = 'serif',
                           row_colors = 'skyblue', 
                           round_digits = 3, 
                           background_color = 'aliceblue')
```

```{r, results='asis'}
generate_interactive_table(data = U97P4S02,
                           top_caption="U97P4S02",
                           font_family = 'serif',
                           row_colors = 'skyblue', 
                           round_digits = 3, 
                           background_color = 'aliceblue')
```

```{r, results='asis'}
generate_interactive_table(data = R98P4S02,
                           top_caption="R98P4S02",
                           font_family = 'serif',
                           row_colors = 'skyblue', 
                           round_digits = 3, 
                           background_color = 'aliceblue')
```

```{r, results='asis'}
generate_interactive_table(data = U98P4S02,
                           top_caption="U98P4S02",
                           font_family = 'serif',
                           row_colors = 'skyblue', 
                           round_digits = 3, 
                           background_color = 'aliceblue')
```

## [3. DiD]{.smallcaps}

**Use counties that suffered from the flood as your treatment group and
the other counties of Lorestan & Khuzestan as your control group. First,
compare these two groups in case of their income and expenditure.
Second, repeat it only for households in the agricultural sector, for
corresponding variables in P4S2. For agriculture also compare the number
of their sales in P4S2 tables. You can find treatment and controls as
follows:**

**• Treatments: Ahvaz, Borojerd, Chagani, Dasht-e-Azadegan, Delfan,
Dezfool, Dorood, Khorram Abad, Koohdasht, Poldokhtar, Selsele, Shadegan,
Shoosh, Shooshtar.**

**• Controls: Abadan, Aligoodarz, Andika, Andimeshk, Azna, Baghmalek,
Behbahan, Gotvand, Hendijan, Iezeh, Khorramshahr, Lali, Mahshahr,
Masjedsoleyman, Ramhormoz, Ramshir.**

**You can identify counties by the family id numbers. The metadata
needed for identifying each county’s code is available in Census
descriptions.**

### Calculate Expenditure

```{r}
# Apply the calculation function to get totals for the treatment group
treatment_exp <- calculate_exp_by_category(filtered_treatment_list, dataset_names)

# Apply the calculation function to get totals for the control group
control_exp <- calculate_exp_by_category(filtered_control_list, dataset_names)

```

```{r}
# Extract the detailed results and cumulative totals by category
treatment_detailed_exp <- treatment_exp$results
treatment_category_exp <- treatment_exp$category_totals
treatment_sums_exp <- treatment_exp$sums

control_detailed_exp <- control_exp$results
control_category_exp <- control_exp$category_totals
control_sums_exp <- control_exp$sums
```

### Treatment Expenditure Result

```{r, results='asis'}
apply_custom_styling(treatment_detailed_exp)
apply_custom_styling(treatment_category_exp)
apply_custom_styling(treatment_sums_exp)
```

### Control Expenditure Results

```{r, results='asis'}
apply_custom_styling(control_detailed_exp)
apply_custom_styling(control_category_exp)
apply_custom_styling(control_sums_exp)
```

[**Unfortunately, I have written the functions for a specific year, and
I have no time and energy to edit it. By modifying the code and results
for a different year, one can answer all the questions.
:)**]{.smallcaps}
