---
title: "Metrics II - Homework 7"
author: "Farnam Taheri"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: readable
    highlight: kate
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
    code_folding: hide
    css: styles.css
    includes:
      before_body: navbar.html
editor_options:
  markdown:
    wrap: 72
---

```{r setup, warning=FALSE, message=FALSE}
# Setup chunk to load required packages and set global options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)

# Load necessary libraries
library(DT)          # Table display
library(lubridate)   # Date handling
library(dplyr)       # Data manipulation
library(ggplot2)     # Data visualization
library(plm)         # Panel data models
library(lmtest)      # Diagnostic tests for linear models
library(sandwich)    # Robust standard errors
library(stargazer)   # To create regression tables
library(plotly)      # Interactive visulization
library(lme4)
library(haven)
library(tidyverse)
library(knitr)
```

# [Part I: The COVID-19 Pandemic and the Effectiveness of Lockdown]{.smallcaps} {#part1}

**Consider the following model:**
$$covid\_cases_{it} = β_0 + β_1 stringency\_index_{it} + ϵ_{it} (1)$$
**where the variable** $covid_cases_{it}$ **measures pandemic situation
in the** $i$**-th country,** $t$ **is the time period,**
$stringency\_index_{it}$ **is the so-called stringency index and it
measures policy responses that governments have taken to react the
COVID-19 pandemic (higher value is related to a larger number of
policies or stricter policies) and** $ϵ_{it}$ **is the error term.**

## [Preliminaries]{.smallcaps}

### Define the Required Functions

```{r}
hausman_test <- function(fe_model, re_model) {
  # Conduct the Hausman test
  hausman_test <- phtest(fe_model, re_model)
  
  # Round the p-value to four digits
  rounded_p_value <- round(hausman_test$p.value, 4)
  
  # Determine interpretation based on the p-value
  if (hausman_test$p.value < 0.05) {
    interpretation <- paste(
      "The Hausman test indicates a significant difference between the fixed effects (FE) and random effects (RE) models (p-value =",
      rounded_p_value, 
      "). Given the low p-value (< 0.05), we reject the null hypothesis that the RE model is consistent. This suggests that the FE model is preferred, implying that there are likely unobserved time-invariant variables correlated with the explanatory variables. These unobserved variables may introduce bias in the RE model, making the FE model a more reliable choice for our analysis."
    )
  } else {
    interpretation <- paste(
      "The Hausman test shows no significant difference between the fixed effects (FE) and random effects (RE) models (p-value =",
      rounded_p_value,
      "). With a p-value greater than 0.05, we fail to reject the null hypothesis, indicating that the RE model is not inconsistent. This suggests that either the FE or RE model could be used. However, if there are concerns about unobserved time-invariant variables correlating with the explanatory variables, the FE model might still be preferred to avoid potential biases."
    )
  }
  
  # HTML output
  cat("<p><strong><center>Hausman Test Results</center></strong></p>",
      "<p><strong>P-value:</strong>", rounded_p_value,
      "<br><strong>Interpretation:</strong>", interpretation, "</p>")
}
```

```{r}
bp_test <- function(pooled_model) {

  # Conduct the BP test
  bp_test <- plmtest(pooled_model, type = "bp")
  
  # Round the p-value to four digits
  rounded_p_value <- bp_test$p.value
  
  # Determine interpretation based on the p-value
  if (bp_test$p.value < 0.05) {
    interpretation <- paste(
      "The Breusch-Pagan test indicates significant cross-state heterogeneity (p-value =",
      rounded_p_value,
      "). Given the low p-value (< 0.05), we reject the null hypothesis of no cross-state heterogeneity. This suggests that the random effects model is more appropriate than the pooled OLS model, as it accounts for unobserved effects that vary across states."
    )
  } else {
    interpretation <- paste(
      "The Breusch-Pagan test shows no significant cross-state heterogeneity (p-value =",
      rounded_p_value,
      "). With a p-value greater than 0.05, we fail to reject the null hypothesis, indicating that the pooled OLS model may be sufficient for the data. However, this does not rule out the possibility of heterogeneity; it simply suggests that the test did not detect it at the chosen significance level."
    )
  }
  
  # HTML output
  cat("<p><strong><center>Breusch-Pagan Test Results</center></strong></p>",
      "<p><strong>P-value:</strong>", rounded_p_value,
      "<br><strong>Interpretation:</strong>", interpretation, "</p>")
}

```

```{r}
#' Generate an Interactive Table Using DT
#'
#' This function generates an interactive table using the DT package.
#'
#' @param data A data frame to display.
#' @param caption The caption for the table (default: NULL).
#' @param col_names Column names for the table (default: NULL).
#' @param options List of options for the DT::datatable function (default: list(pageLength = 5, scrollX = TRUE)).
#' @param row_colors Row colors for the table (default: NULL).
#' @param note Note to add below the table (default: NULL).
#' @param footnote Footnote to add below the table (default: NULL).
#' 
#' @return Interactive table.
#' @examples
#' generate_interactive_table(mtcars, caption = "Car Dataset")
generate_interactive_table <- function(data, caption = NULL, col_names = NULL,
                                       options = list(pageLength = 5, scrollX = TRUE),
                                       row_colors = NULL, note = NULL, footnote = NULL) {
  if (!is.data.frame(data)) stop("Input data must be a data frame.")
  
  table <- datatable(data, caption = htmltools::tags$caption(
    style = 'caption-side: top; text-align: center;', caption),
    colnames = col_names, options = options)
  
  if (!is.null(row_colors)) {
    table <- table %>% formatStyle(0, target = 'row', backgroundColor = row_colors)
  }
  
  if (!is.null(note)) {
    table <- table %>% formatStyle(columns = 1:ncol(data), valueColumns = note, fontSize = '80%')
  }
  
  if (!is.null(footnote)) {
    table <- table %>% formatStyle(columns = 1:ncol(data), valueColumns = footnote, fontSize = '80%', italic = TRUE)
  }
  
  return(table)
}

```

### Set the working directory

```{r message=FALSE, warning=FALSE}
# Set the working directory to the location of the data files
setwd("./Data")

# Assign the working directory to a variable for future reference
working_dir <- getwd()
```

### Load Data

First, we download the latest data related to COVID-19 and the
stringency index from the Our World in Data dataset.

```{r message=FALSE, warning=FALSE}
# Read the CSV file from the working directory and Convert to tibble for better readability
covid_data <- as_tibble(read.csv("./Data/owid-covid-data.csv"))
```

```{r}
# Convert 'date' column to Date type
covid_data <- covid_data %>%
  mutate(date = as.Date(date, format="%Y-%m-%d"))

# Aggregate data by month while retaining additional relevant columns
covid_data <- covid_data %>%
  # Create a new column for the month
  mutate(month = floor_date(date, "month")) %>%
  # Group by location and month
  group_by(location, month) %>%
  # Summarize the data by aggregating relevant columns
  summarize(
    new_cases_smoothed = sum(new_cases_smoothed, na.rm = TRUE),
    new_cases = sum(new_cases, na.rm = TRUE),
    positive_rate = mean(positive_rate, na.rm = TRUE),
    tests_per_case = mean(tests_per_case, na.rm = TRUE),
    total_vaccinations = sum(total_vaccinations, na.rm = TRUE),
    people_vaccinated = sum(people_vaccinated, na.rm = TRUE),
    people_fully_vaccinated = sum(people_fully_vaccinated, na.rm = TRUE),
    total_boosters = sum(total_boosters, na.rm = TRUE),
    new_vaccinations_smoothed = sum(new_vaccinations_smoothed, na.rm = TRUE),
    stringency_index = mean(stringency_index, na.rm = TRUE),
    population = first(population),
    population_density = first(population_density),
    median_age = first(median_age),
    aged_65_older = first(aged_65_older),
    aged_70_older = first(aged_70_older),
    gdp_per_capita = first(gdp_per_capita),
    extreme_poverty = first(extreme_poverty),
    cardiovasc_death_rate = first(cardiovasc_death_rate),
    diabetes_prevalence = first(diabetes_prevalence),
    handwashing_facilities = first(handwashing_facilities),
    hospital_beds_per_thousand = first(hospital_beds_per_thousand),
    life_expectancy = first(life_expectancy),
    human_development_index = first(human_development_index),
    excess_mortality = sum(excess_mortality, na.rm = TRUE),
    excess_mortality_cumulative = sum(excess_mortality_cumulative, na.rm = TRUE),
    excess_mortality_cumulative_absolute = sum(excess_mortality_cumulative_absolute, na.rm = TRUE),
    excess_mortality_cumulative_per_million = sum(excess_mortality_cumulative_per_million, na.rm = TRUE)
  ) %>%
  # Ungroup the data frame
  ungroup()

# Add a 'month' factor column for fixed effects
covid_data <- covid_data %>%
  mutate(month_factor = factor(month(month)))
```

```{r}
# Display the first few rows of the dataset to understand its structure
set.seed(42)    # Set seed for reproducibility
covid_random_rows <- sample_n(covid_data, 250) # Pick 50 random rows from the dataset
generate_interactive_table(covid_random_rows,
                           col_names=colnames(covid_random_rows),
                           row_colors = "aliceblue")
```

## [A. Expected Sign of]{.smallcaps} $β_1$

**What signs would you expect on the coefficient** $β_1$**. Why?**

The coefficient $β_1$ measures the relationship between the stringency
index (policy responses) and the number of COVID-19 cases. We expect
$β_1$ to be negative because stricter policies (higher stringency index)
should lead to fewer COVID-19 cases. Stricter policies are intended to
reduce the spread of the virus by limiting social interactions and
mobility. Thus a higher stringency (stricter policies) should ideally
lead to fewer cases due to reduced transmission.

## [B. Choose Proxy Variable]{.smallcaps}

**Download the latest data related to the COVID-19 pandemic and the
stringency index. You can use the Our World in Data dataset ([link to
the csv.
file](https://covid.ourworldindata.org/data/owid-covid-data.csv "Covid Dataset"),
[description of
variables](https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-codebook.csv "Variables' description")).
Choose proxy variable for** $covid\_cases_{it}$ **and discuss this
choice.**

Selecting the right proxy variable for $covid\_cases_{it}$ in Model (1)
is crucial for understanding the evolving COVID-19 pandemic. Among
available options, the `new_cases` metric stands out as a reliable
choice. It directly reflects the current state of the pandemic in each
country at any given time, making it a fundamental indicator of the
virus's impact on populations. Integrating `new_cases` into the model
provides timely insights into the pandemic's dynamics, essential for
evaluating government responses.

While `new_cases` is a good choice, comparing it with the
`new_cases_smoothed` variable reveals some nuances. `New_cases_smoothed`
offers benefits like stabilizing volatile data and addressing reporting
errors through a 7-day smoothing process. This creates a more consistent
trend in COVID-19 cases over time, facilitating comparisons across
countries and analyzing the relationship with policy stringency.

However, despite these advantages, the preference leans towards using
the raw `new_cases` metric over `new_cases_smoothed.` This decision
arises from concerns about potential autocorrelation when aggregating
smoothed data on a monthly basis. Autocorrelation can bias estimates and
inflate errors, especially when data is aggregated more frequently than
originally collected. Opting for the raw `new_cases` metric helps avoid
these issues, ensuring the analysis remains accurate and reliable.

In conclusion, while both `new_cases` and `new_cases_smoothed` have
their merits, the raw `new_cases` metric is the preferred choice for
Model (1).

## [C. Model Comparison]{.smallcaps}

**Estimate the parameter of (1) using pooled regression as well as the
RE and FE estimators. Remember about clustered standard errors. Are the
estimates of** $β_1$ **in line with your expectations? Test the
heterogeneity in the RE and FE models. Compare RE and FE models with the
Hausman test.**

#### Model Estimation

First, we will estimate different models and analyze the models'
results.

```{r, results=FALSE}
# Convert data to panel data frame
  covid_pdata <- pdata.frame(covid_data, index = c("location", "month"))
  
# Estimate the extended model using pooled OLS
pooled_model <- plm(
  formula = new_cases ~ stringency_index,  # Specify the model formula
  data = covid_pdata,                     # Provide the data
  model = "pooling"                      # Use the pooling model (Pooled OLS)
)
  
# Estimate the extended model using pooled OLS
re_model <- plm(
  formula = new_cases ~ stringency_index,
  data = covid_pdata,                     
  model = "random"                      # Use the random effect model 
)
  
  
  
# Estimate the extended model using pooled OLS
fe_model <- plm(
  formula = new_cases ~ stringency_index,
  data = covid_pdata,                  
  model = "within"                      # Use the fixed effect model 
)
  
# Compute clustered standard errors
pooled_se <- coeftest(pooled_model,                   # Compute coefficient test
                  vcov = vcovHC(                      # Using the robust covariance estimator
                            pooled_model,             # Based on the pooled model
                            type = "HC1",             # Type of robust covariance estimator
                            cluster = "group")     # Clustering by country
                      )

re_se <- coeftest(re_model,
              vcov = vcovHC(re_model, type = "HC1", cluster = "group")
                           )
fe_se <- coeftest(fe_model,
              vcov = vcovHC(fe_model, type = "HC1", cluster = "group")
                           )


# Display results using stargazer
stargazer_models <- stargazer( 
  pooled_model, re_model, fe_model,                                    # Models to display
  type = "html",                                                       # Output type as HTML
  title = "<center>COVID-19 Cases and Stringency Index<center>",       # Title for the table
  dep.var.labels = c("New COVID-19 Cases"),                            # Dependent variable label
  column.labels = c("Pooled OLS", "Random Effects", "Fixed Effects"),  # Model names for columns
  
  covariate.labels = c(              # Labels for covariates
                       "Constant"="Constant",  
                       "stringency_index"="Stringency Index",
                       "population_density"="Population Density",
                       "median_age"="Median Age",
                       "month_factor"="Month Factor"), 

  digits = 2,                        # Number of digits to display
  align = TRUE,                      # Align columns
  style = "AER",                     # Use AER style
  intercept.bottom = FALSE,          # Display intercepts at the top
  intercept.top = TRUE,              # Display intercepts at the bottom
  se = list(                         # Clustered standard errors
    pooled_se[, 2], re_se[, 2], fe_se[, 2]
    ) 
)
```

```{r echo=FALSE, results='asis'}
cat("<center>",stargazer_models,"<center>")
```

#### Heterogeneity Test

Now we will perform the Hausman test in oreder tocompare **fixed
effect** and **random** **effect** models.

```{r, results='asis'}
# Conducting the Hausman test utilizing the defined function
hausman_test(fe_model, re_model)
```

## [D. Models with Control Varialbes]{.smallcaps}

**Choose at least two control variables that potentially affect the
outcome variable and extend (1) by the selected variables. What sign
would you expect on the related coefficients? Estimate the extended
regression with pooled regression as well as the RE and FE estimators.
Are the estimates of** $β_1$ **in line with your expectations? Test the
heterogeneity in the RE and FE models. Compare RE and FE models with the
Hausman test.**

### Model Estimation

First, we will estimate different models and analyze the models'
results.

```{r, results=FALSE}

# Convert data to panel data frame with the new control variables included
  covid_pdata <- pdata.frame(covid_data, index = c("location", "month"))
  
  # Estimate the extended model using pooled OLS
  pooled_model_extended <- plm(
    formula = new_cases_smoothed ~ stringency_index + population_density + gdp_per_capita + median_age,  # Specify the model formula
    data = covid_pdata,                     # Provide the data
    model = "pooling"                      # Use the pooling model (Pooled OLS)
  )
  
  # Estimate the extended model using pooled OLS
  re_model_extended <- plm(
    formula = new_cases_smoothed ~ stringency_index + population_density + gdp_per_capita + median_age,  
    data = covid_pdata,                     
    model = "random"                      # Use the random effect model 
  )
  
  # Estimate the extended model using pooled OLS
  fe_model_extended <- plm(
    formula = new_cases_smoothed ~ stringency_index + population_density + gdp_per_capita + median_age,
    data = covid_pdata,                  
    model = "within"                      # Use the fixed effect model 
  )
  
# Compute clustered standard errors
pooled_se_extended <- coeftest(pooled_model_extended,
                           vcov = vcovHC(pooled_model_extended, type = "HC1", cluster = "group")
)
re_se_extended <- coeftest(re_model_extended,
                           vcov = vcovHC(re_model_extended, type = "HC1", cluster = "group")
                           )
fe_se_extended <- coeftest(fe_model_extended,
                           vcov = vcovHC(fe_model_extended, type = "HC1", cluster = "group")
                           )
 
# Display results using stargazer
stargazer_models_extended <- stargazer(
  pooled_model_extended, re_model_extended, fe_model_extended,
  
  title = "<center>COVID-19 Cases and Stringency Index<center>", 
  dep.var.labels = c("New COVID-19 Cases"),
  column.labels = c("Pooled OLS", "Random Effects", "Fixed Effects"),

  covariate.labels = c("Constant"="Constant",
                       "stringency_index"="Stringency Index",
                       "population_density"="Population Density",
                       "gdp_per_capita"="GDP per Capita",
                       "median_age"="Median Age",
                       "month_factor"="Month Factor"), 
  
  type = "html", digits = 2, align = TRUE, style = "AER",
  intercept.bottom = FALSE, intercept.top = TRUE,
  se = list(pooled_se_extended[, 2], re_se_extended[, 2], fe_se_extended[, 2])
)
```

```{r echo=FALSE, results='asis'}
cat("<center>",stargazer_models_extended,"<center>")
```

### Heterogeneity Test

Now we will perform the Hausman test in oreder tocompare **fixed
effect** and **random** **effect** models.

```{r, results='asis'}
# Conducting the Hausman test utilizing the defined function
hausman_test(fe_model_extended, re_model_extended)
```

## [E. Time Effect]{.smallcaps}

**Add time effects to the previous regressions and re-estimate
parameters. Is there any trend effect? If so, try to visualize this
effect for all models. Once again, are the estimates of** $β_1$ **in
line with your expectations? Test the heterogeneity in the RE and FE
models. Compare RE and FE models with the Hausman test.**

### Model Estimation

```{r, results=FALSE}
# Estimate the extended model with time effects using pooled OLS
pooled_model_time <- plm(
  formula = new_cases ~ stringency_index + population_density + gdp_per_capita + factor(month),
  data = covid_pdata,
  model = "pooling"
)

# Estimate the extended model with time effects using fixed effects
fe_model_time <- plm(
  formula = new_cases ~ stringency_index + population_density + gdp_per_capita + factor(month),
  data = covid_pdata,
  model = "within"
)

# Estimate the extended model with time effects using random effects with lme4
re_model_time <- lmer(
  formula = new_cases ~ stringency_index + population_density + gdp_per_capita + factor(month) + (1 | location),
  data = covid_data
)

# Compute clustered standard errors
pooled_se_time <- coeftest(pooled_model_time, 
                           vcov = vcovHC(pooled_model_time, type = "HC1", cluster = "group")
)

fe_se_time <- coeftest(fe_model_time, 
                       vcov = vcovHC(fe_model_time, type = "HC1", cluster = "group")
)


# Display results using stargazer
stargazer_models_time <- stargazer(
  pooled_model_time, fe_model_time, re_model_time,
  title = "<center>Extended Models with Time Effects: COVID-19 Cases and Stringency Index<center>", 
  dep.var.labels = c("New COVID-19 Cases"), 
  column.labels = c("Pooled OLS", "Fixed Effects", "Random Effects"), 
  covariate.labels = c("Constant", "Stringency Index", "Population Density", "GDP per Capita", "Month Factor"), 
  type = "html", digits = 2, align = TRUE, style = "AER", 
  intercept.bottom = FALSE, intercept.top = TRUE,
  se = list(pooled_se_time[, 2], fe_se_time[, 2])
)
```

```{r echo=FALSE, results='asis'}
cat("<center>",stargazer_models_time,"<center>")
```

## [F. Discussion of Limitations and Potential Extensions]{.smallcaps}

### **Limitations:**

1.  **Omitted Variable Bias:** The model may not account for all
    relevant factors that influence COVID-19 cases. For instance,
    variables like healthcare system quality, public health literacy,
    and international travel could significantly impact the spread of
    the virus but may not be included in the analysis.

2.  **Measurement Error:** The data on COVID-19 cases and stringency
    index might suffer from reporting errors. Different countries have
    varying testing capacities and reporting standards, which can lead
    to underreporting or overreporting of cases.

3.  **Cross-sectional Dependence:** The assumption that observations
    across countries are independent might be violated. Countries may
    influence each other's pandemic response or share common shocks,
    such as new virus variants, which can lead to spurious results.

4.  **Nonlinearity and Threshold Effects:** The relationship between
    stringency measures and COVID-19 cases might be nonlinear. For
    example, there could be threshold effects where stringency measures
    only become effective after surpassing a certain level.

5.  **Endogeneity:** There may be reverse causality where the number of
    COVID-19 cases influences policy stringency. Additionally,
    unobserved factors could affect both the stringency index and
    COVID-19 cases, leading to biased estimates.

    **Reverse Causality:**

    -   If a government increases the stringency of its policies in
        response to a rise in COVID-19 cases, then the number of cases
        is influencing the policy response. This creates a situation
        where the cause and effect are influencing each other, making it
        difficult to determine the direction of causality.

    **Unobserved Factors:**

    -   The effectiveness of stringent policies might be influenced by
        public behavior and compliance, which are not directly observed
        or measured. If people voluntarily engage in social distancing
        and reduce mobility even before official policies are
        implemented, the stringency index might not capture the actual
        level of social distancing taking place.

6.  **Temporal Dynamics:** The model may not capture the dynamic nature
    of the pandemic. The effects of policies on COVID-19 cases can have
    time lags, and the impact may vary over different stages of the
    pandemic.

### **Potential Extensions:**

1.  **Instrumental Variables (IV) Approach:** To address potential
    endogeneity, an IV approach could be used where instruments are
    variables that are correlated with the stringency index but not
    directly with COVID-19 cases.

2.  **Dynamic Panel Data Models:** Using lag models can help control for
    unobserved heterogeneity and capture the lagged effects of policies.

3.  **Spatial Econometric Models:** These models can account for
    cross-sectional dependence and spillover effects between countries,
    providing a more accurate representation of the pandemic's spread.

4.  **Robustness Checks:** Conducting robustness checks using different
    model specifications, time periods, and subsets of data can help
    validate the findings.

# [Part II: Demand for Cigarettes]{.smallcaps} {#part2}

Baltagi and Li (1986) study a demand for cigarettes. Consider the subset
of their dataset ([link to the dta.
file](Cigarettes'%20Dataset "https://drive.google.com/file/d/1x5EDFL4JW72QdNuxZY_Kx7eSmgbFLhNR/view"))
that contains data on 46 US states from 1963-1982. The following
variables are available:

• $state_{it}$ – the identification number of state

• $year_{it}$ – the year,

• $price_{it}$– price per pack of cigarettes (in nominal terms),

• $cpi_{it}$ – consumer price index (1983=100),

• $income_{it}$– per capita disposable income (in nominal terms),

• $sales_{it}$ – cigarette sales in packs per capita.

## [Preliminaries]{.smallcaps}

```{r}
# Load the dataset
cig_data <- as_tibble(read_dta("./Data/Cigar.dta"))

# Construct variables
cig_data <- cig_data %>%
  mutate(
    ln_sales = log(sales),
    ln_income = log(income / cpi * 100),
    ln_price = log(price / cpi * 100)
  )
```

## [A. Log-transformation]{.smallcaps}

**At the first stage, consider the following econometric model
describing demand for cigarettes:**

$$\ln C_{it} = β_0 + β_1 ln Y_{it} + β_2 ln P_{it} + ϵ_{it} \quad (2)$$
**where** $C_{it}$ is the consumption of cigarettes, $Y_{it}$ is the
real disposable income per capita and $P_{it}$ is the real price per
pack of cigarettes. Why the log transformation is useful here?

The log transformation is useful for several reasons:

1.  **Linearity**: The relationship between cigarette consumption,
    income, and price may not be linear. By using a log-log model, we
    can linearize a potentially multiplicative relationship. This means
    that percentage changes in income or price are associated with
    percentage changes in cigarette consumption, which can be modeled
    using linear regression.

2.  **Elasticity Interpretation**: In this context, the coefficients
    $\beta_1$ and $\beta_2$ directly provide the elasticities of demand.
    The income elasticity ($\beta_1$) indicates the percentage change in
    cigarette consumption for a one percent change in real disposable
    income per capita. Similarly, the price elasticity ($\beta_2$)
    represents the percentage change in cigarette consumption for a one
    percent change in the real price per pack of cigarettes. This is
    particularly useful for policymakers and economists who are
    interested in understanding how changes in economic factors affect
    cigarette consumption.

3.  **Heteroscedasticity**: Economic data often exhibit
    heteroscedasticity, where the variance of the error term changes
    across observations. Log transformation can mitigate this issue,
    leading to more efficient and reliable estimates, as the variance of
    the percentage changes is more likely to be constant.

4.  **Normality**: The distribution of economic data can be skewed, with
    a few large values. Log transformation helps in achieving a
    distribution that is closer to normal, which is a key assumption in
    many statistical methods, including regression analysis.

5.  **Outliers**: Economic datasets can have outliers due to data entry
    errors, unusual events, or structural changes. Log transformation
    reduces the impact of outliers because it compresses the scale,
    bringing the data points closer together and making the dataset more
    homogenous.

## [B. Appropriate Variables]{.smallcaps}

**Based on the available dataset, propose and discuss the construction
of variables that are able to identify empirically** $C_{it}$, $Y_{it}$
and $P_{it}$.

-   **Cigarette Consumption**: This variable can be identified using
    `$sales_{it}$`, which represents cigarette sales in packs per
    capita. Since the econometric model uses the natural logarithm of
    cigarette consumption, we would take the log of `salesit` to
    get$\ln(C_{it})$.

-   **Real Disposable Income Per Capita**: The `$income_{it}$` variable
    indicates per capita disposable income in nominal terms. To adjust
    for inflation and convert it into real terms, we divide
    `$income_{it}$` by `$cpi_{it}$` (Consumer Price Index) and multiply
    by 100 (to reference the base year 1983=100). The natural logarithm
    of this adjusted income gives us $\ln(Y\_{it})$.

-   $P_{it}$ (Real Price Per Pack of Cigarettes): Similarly,
    `$price_{it}$` provides the price per pack of cigarettes in nominal
    terms. We adjust it for inflation by dividing `$price_{it}$` by
    `$cpi_{it}$` and multiplying by 100. Taking the natural logarithm of
    this adjusted price yields $\ln(P_{it})$.

By transforming these variables into their logarithmic forms, we can
analyze the elasticities and interpret the coefficients in the
regression model more meaningfully.

## [C. Elasticity Expectations]{.smallcaps}

For the coefficients $\beta_1$ and $\beta_2$ in the demand equation:

$$
\ln(C_{it}) = \beta_0 + \beta_1 \ln(Y_{it}) + \beta_2 \ln(P_{it}) + \epsilon_{it}
$$

-   **Expectation for** $\beta_1$ (Income Elasticity): We would
    generally expect a positive sign for \$\beta\_1 \$, as income
    elasticity of demand for cigarettes is typically positive. This
    means that as real disposable income increases, people can afford to
    buy more cigarettes, leading to an increase in cigarette
    consumption.

-   **Expectation for** $\beta_2$ (Price Elasticity): Typically, the
    sign for $\beta_2$ is expected to be negative, aligning with the law
    of demand which posits that as the price of a good increases, the
    quantity demanded decreases, all else being equal. Consequently, a
    rise in the real price of cigarettes should result in reduced
    cigarette consumption.

However, it's important to note that if cigarettes are a Giffen good,
the sign for $\beta_2$ could actually be positive. This counterintuitive
scenario suggests that an increase in the real price of cigarettes might
lead to higher cigarette consumption due to the dominant income effect
overriding the substitution effect.

## [D. Model Comparison]{.smallcaps}

**Estimate the model using a pooled OLS, a RE estimator and FE
estimator. Report all estimates in one table, interpret (remember that
the variables are in logs) and discuss your results. Are they as you
expected? Explain why these estimates differ.**

### Model Estimation

```{r, results=FALSE}
# Model estimation
cig_pooled_model <- plm(
  formula = ln_sales ~ ln_income + ln_price,
  data = cig_data,
  model = "pooling"
)

cig_fe_model <- plm(
  formula = ln_sales ~ ln_income + ln_price,
  data = cig_data,
  model = "within"
)

cig_re_model <- plm(
  formula = ln_sales ~ ln_income + ln_price,
  data = cig_data,
  model = "random"
)

# Compute standard errors
cig_pooled_se <- coeftest(cig_pooled_model, vcov = vcovHC(cig_pooled_model, type = "HC1", cluster = "group"))
cig_fe_se <- coeftest(cig_fe_model, vcov = vcovHC(cig_fe_model, type = "HC1", cluster = "group"))
cig_re_se <- coeftest(cig_re_model, vcov = vcovHC(cig_re_model, type = "HC1", cluster = "group"))

# Display results using stargazer
stargazer_models <- stargazer(
  cig_pooled_model, cig_fe_model, cig_re_model,
  title = "<center>Models of Cigarette Consumption<center>", 
  dep.var.labels = c("Log of Cigarette Sales"), 
  column.labels = c("Pooled OLS", "Fixed Effects", "Random Effects"), 
  covariate.labels = c("Constant", "Log of Income", "Log of Price"), 
  type = "html", digits = 2, align = TRUE, style = "AER", 
  intercept.bottom = FALSE, intercept.top = TRUE,
  se = list(cig_pooled_se[, 2], cig_fe_se[, 2], cig_re_se[, 2])
)
```

```{r echo=FALSE, results='asis'}
# Output the HTML code for the table
cat("<center>", stargazer_models, "<center>")

```

### Result Analysis

-   **Pooled OLS**: The constant term is significant and positive,
    suggesting a base level of cigarette sales regardless of income or
    price. The positive coefficient for income indicates that cigarette
    sales increase as income increases. The negative price elasticity
    suggests that sales decrease as prices rise, which is consistent
    with the law of demand.

-   **Fixed Effects**: The income coefficient is not significant,
    indicating that within entities (such as individuals or regions),
    changes in income do not significantly affect cigarette sales. The
    negative price elasticity is significant and slightly less in
    absolute value than in the pooled OLS, suggesting that within
    entities, consumers are less responsive to price changes.

-   **Random Effects**: The constant term is significant and higher than
    in the pooled OLS. The income coefficient is again insignificant,
    and the price elasticity is the same as in the fixed effects model,
    suggesting that when accounting for random differences across
    entities, income does not have a significant effect on cigarette
    sales, but price changes do.

The results might differ from expectations if cigarettes are a Giffen
good, as discussed earlier, or if there are other factors affecting
cigarette consumption that are not included in the model. The lack of
significance for income in the fixed and random effects models suggests
that other factors, possibly cultural or policy-related, may play a more
substantial role in determining cigarette consumption than income alone.

## [E. Lagrange Multiplier]{.smallcaps}

**Test the hypothesis that there are no cross-state differences and test
the presence of the cross-state heterogeneity using the Lagrange
multiplier statistic. Do these results lead to the same conclusions?**

```{r, results='asis'}
bp_test(cig_pooled_model)
```

## [F. Hausman Test]{.smallcaps}

**Perform Hausman’s test for the estimates obtained in (d) for the RE
and FE model. Interpret this result.**

## [G. Year Dummies]{.smallcaps}

```{r, results=FALSE}
# Convert the 'year' column to a factor with levels for each year
cig_data$year <- factor(cig_data$year)

# Use the 'model.matrix()' function to create dummy variables for each year
# The base level (first year) is automatically excluded to prevent multicollinearity
year_dummies <- model.matrix(~ year - 1, data = cig_data)

# Bind the year dummy variables to the original dataset
cig_data <- cbind(cig_data, year_dummies)

# Re-estimate the FE model with year dummies
cig_fe_model_year <- plm(
  formula = ln_sales ~ ln_income + ln_price + year,
  data = cig_data,
  model = "within"
)

# Re-estimate the RE model with year dummies
cig_re_model_year <- plm(
  formula = ln_sales ~ ln_income + ln_price + year,
  data = cig_data,
  model = "random"
)

# Re-estimate the pooled OLS model with year dummies
cig_pooled_model_year <- lm(
  formula = ln_sales ~ ln_income + ln_price + year,
  data = cig_data
)

# Compute standard errors for the pooled model with year dummies
cig_pooled_se_year <- coeftest(cig_pooled_model_year, vcov = vcovHC(cig_pooled_model_year, type = "HC1", cluster = "group"))

# Compute standard errors for the new models
cig_fe_se_year <- coeftest(cig_fe_model_year, vcov = vcovHC(cig_fe_model_year, type = "HC1", cluster = "group"))
cig_re_se_year <- coeftest(cig_re_model_year, vcov = vcovHC(cig_re_model_year, type = "HC1", cluster = "group"))


# Display results using stargazer
stargazer_models <- stargazer(
  cig_pooled_model_year, cig_fe_model_year, cig_re_model_year,
  title = "<center>Models of Cigarette Consumption with Year Dummies<center>", 
  dep.var.labels = c("Log of Cigarette Sales"), 
  column.labels = c("Pooled OLS with Year Dummies", "Fixed Effects with Year Dummies", "Random Effects with Year Dummies"), 
  covariate.labels = c("Constant", "Log of Income", "Log of Price", "Year Dummies"), 
  type = "html", digits = 2, align = TRUE, style = "AER", 
  intercept.bottom = FALSE, intercept.top = TRUE,
  se = list(cig_pooled_se_year[, 2], cig_fe_se_year[, 2], cig_re_se_year[, 2])
)
```

```{r echo=FALSE, results='asis'}
# Output the HTML code for the table
cat("<center>", stargazer_models, "<center>")
```

## [H. Hausman Test]{.smallcaps}

**Perform Hausman’s test for the estimates obtained in (g)? Discuss this
result.**

```{r results='asis'}
hausman_test(cig_fe_model_year, cig_re_model_year)
```
