---
title: 'Metrics II - Homework 9'
author: 'Farnam Taheri'
date: '`r Sys.Date()`'
output:
  html_document:
    theme: readable
    highlight: kate
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
    code_folding: hide
    css: styles.css
    includes:
      before_body: navbar.html
editor_options:
  markdown:
    wrap: 72
---

# [Part I: Angrist and Lavy (1999)]{.smallcaps} {#part1}

**In this part, you are asked to replicate some of the results of
Angrist and Lavy (1999) using the provided data file `final5.dta`.**

## **Variable Description**

-   **`schlcode`: School code**
-   **`csize`: School size (to be renamed to `enroll`)**
-   **`classize`: Class size**
-   **`tpiuach`: Percent disadvantaged (to be renamed to `pdis`)**
-   **`avgverb`: Average verbal**
-   **`avgmath`: Average math**
-   **`verbsize`: Reading size**
-   **`mathsize`: Math size**

## Preliminaries

### Load Necessary Libraries

```{r setup, warning=FALSE, message=FALSE, results=FALSE}
# Setup chunk to load required packages and set global options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)

# Install necessary packages if not already installed
if (!requireNamespace('tidyverse', quietly = TRUE)) install.packages('tidyverse')
if (!requireNamespace('foreign', quietly = TRUE)) install.packages('foreign')
if (!requireNamespace('DT', quietly = TRUE)) install.packages('DT')
if (!requireNamespace('htmltools', quietly = TRUE)) install.packages('htmltools')
if (!requireNamespace('stargazer', quietly = TRUE)) install.packages('stargazer')
if (!requireNamespace('AER', quietly = TRUE)) install.packages('AER')
if (!requireNamespace('dplyr', quietly = TRUE)) install.packages('dplyr')

# Load the packages
library(tidyverse)
library(foreign)
library(AER)
library(DT)
library(htmltools)
library(stargazer)
library(dplyr)
```

### Load Data

```{r}
setwd('.')
final5 <- as_tibble(read.dta('./Data/final5.dta'))
```

# [Part II: Tasks]{.smallcaps} {#part2}

## [**1. Assumptions:**]{.smallcaps}

**List the assumptions necessary for consistent estimates in the
regression discontinuity approach.**

To achieve consistent estimates in a regression discontinuity design
(RDD), the following assumptions must be met:

1.  **Continuity at the Cutoff**: The potential outcomes must be
    continuous at the threshold, ensuring no abrupt jumps in the outcome
    variable other than the treatment effect.

2.  **No Manipulation around the Cutoff**: The running variable should
    not be subject to manipulation by individuals or entities that could
    affect treatment assignment.

3.  **Local Randomization**: Units just above and below the cutoff are
    assumed to be similar in both observed and unobserved
    characteristics, mimicking a randomized experiment.

4.  **Correct Functional Form**: The correct functional form of the
    relationship between the running variable and the outcome must be
    specified. This includes modeling the underlying trend accurately.

## [**2. Data Cleaning:**]{.smallcaps}

-   **Rename `c_size` to `enroll` and `tpiuach` to `pdis`.**
-   **Remove school sizes less than 5 students.**
-   **Keep observations with school sizes less than 44 students.**
-   **Retain only the variables listed in the table above.**
-   **Correct any verbal and math grades higher than 100 by subtracting
    100 points.**
-   **Replace class sizes of zero or missing with a `missing value`.**
-   **Report the minimum, maximum, and mean of each variable (summary
    statistics).**

```{r, results=FALSE}
# Assuming final5 is already loaded as a data frame

# Renaming columns
final5 <- final5 %>%
  rename(enroll = c_size, pdis = tipuach)

# Selecting specific columns
final5 <- final5 %>%
  select(schlcode, enroll, classize, pdis, avgverb, avgmath, verbsize, mathsize)

# Filtering rows based on enroll and classize
final5 <- final5 %>%
  filter(enroll > 4 & classize < 45)

# Correcting grades higher than 100
final5 <- final5 %>%
  mutate(
    avgverb = ifelse(avgverb > 100, avgverb - 100, avgverb),
    avgmath = ifelse(avgmath > 100, avgmath - 100, avgmath)
  )

# Replacing zeros with NA
final5 <- final5 %>%
  mutate(
    mathsize = ifelse(mathsize == 0, NA, mathsize),  
    avgmath = ifelse(avgmath == 0, NA, avgmath),    
    verbsize = ifelse(verbsize == 0, NA, verbsize), 
    avgverb = ifelse(avgverb == 0, NA, avgverb)     
  )


descriptive_stats <-  stargazer(as.data.frame(subset(final5, select = -schlcode)),  
          type = 'html',
          title = '<center>Descriptive Statistics<center>',
          label = 'tab:descriptive_stats',
          digits = 2,
          header = FALSE,
          covariate.labels = c('School Code',
                               'Enrollment',
                               'Class Size',
                               'Percent Disadvantaged',
                               'Average Verbal Score',
                               'Average Math Score',
                               'Reading Size',
                               'Math Size'),
          summary.stat = c('min', 'max', 'mean', 'sd', 'p25', 'median', 'p75' ),
          notes = c('Source: Angrist and Lavy (1999)'),
          notes.align = 'l',
          align = TRUE,
          font.size = 'small',
          initial.zero = TRUE,
          style = 'AER')
```

```{r, results='asis'}
# Customize the HTML output with CSS 
descriptive_stats <- descriptive_stats %>%
  gsub("<table", "<table style='width: 100%; border-spacing: 15px; border-collapse: separate; text-align: center;'", .) 

# Print summary statistics table
cat('<center>', descriptive_stats, '<center>')

```

## [**3. Math Grades Estimation:**]{.smallcaps}

-   **Rebuild table 2 for math grades.**
-   **For 2SLS estimation, use**$m = \frac{\text{enroll}}{\text{int}[\frac{\text{enroll}}{40}]+ 1} .$
-   **Handle standard errors appropriately.**

```{r, results=FALSE}
# Create the instrumental variable
final5 <- final5 %>%
  mutate(m = enroll / (floor((enroll - 1) / 40) + 1)) %>%
  mutate(enrollsquared = enroll^2 / 100)
```


```{r, results=FALSE}
# Math Grades Estimation

# OLS estimations
ols_math1 <- lm(avgmath ~ classize, data = final5, na.action = na.exclude)
ols_math2 <- lm(avgmath ~ classize + pdis, data = final5, na.action = na.exclude)
ols_math3 <- lm(avgmath ~ classize + pdis + enroll, data = final5, na.action = na.exclude)

# 2SLS estimations
tsls_math1 <- ivreg(avgmath ~ classize + pdis + enroll | m + pdis + enroll, data = final5, na.action = na.exclude)
tsls_math2 <- ivreg(avgmath ~ classize + pdis + enroll + enrollsquared | m + pdis + enroll + enrollsquared, data = final5, na.action = na.exclude)

# Compute robust standard errors clustered by school code
robust_se_ols1 <- sqrt(diag(vcovCL(ols_math1, cluster = ~schlcode)))
robust_se_ols2 <- sqrt(diag(vcovCL(ols_math2, cluster = ~schlcode)))
robust_se_ols3 <- sqrt(diag(vcovCL(ols_math3, cluster = ~schlcode)))
robust_se_tsls1 <- sqrt(diag(vcovCL(tsls_math1, cluster = ~schlcode)))
robust_se_tsls2 <- sqrt(diag(vcovCL(tsls_math2, cluster = ~schlcode)))

# Generate the table with stargazer
stargazer_math <- stargazer(
  ols_math1, ols_math2, ols_math3, tsls_math1, tsls_math2,
  se = list(robust_se_ols1, robust_se_ols2, robust_se_ols3, robust_se_tsls1, robust_se_tsls2),
  type = 'html',
  title = '<center>Effects of Class Size on Fifth Grade Math Scores<center>',
  label = 'tab:class_size_effects_math',
  digits = 4,
  align = TRUE,
  dep.var.labels = '<center>Math Scores<center>',
  notes = c('Source: Angrist and Lavy (1999)'),
  notes.align = 'l',
  covariate.labels = c('Class size', 'Percent disadvantaged', 'Enrollment', 'Enrollment squared / 100'),
  column.labels = c('OLS', 'OLS', 'OLS', '2SLS', '2SLS'),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("ser", "adj.rsq")
)

```


```{r, results='asis'}
# Display the results in a well-formatted table
cat('<center>', stargazer_math, '<center>')
```

## [**4. Verbal Grades Estimation:**]{.smallcaps}

**Rebuild table 2 for verbal grades.**

```{r, results=FALSE}
# Verbal Scores Estimation
# OLS estimations
ols_verb1 <- lm(avgverb ~ classize, data = final5, na.action = na.exclude)
ols_verb2 <- lm(avgverb ~ classize + pdis, data = final5, na.action = na.exclude)
ols_verb3 <- lm(avgverb ~ classize + pdis + enroll, data = final5, na.action = na.exclude)

# 2SLS estimations
tsls_verb1 <- ivreg(avgverb ~ classize + pdis + enroll | m + pdis + enroll, data = final5, na.action = na.exclude)
tsls_verb2 <- ivreg(avgverb ~ classize + pdis + enroll + enrollsquared | m + pdis + enroll + enrollsquared, data = final5, na.action = na.exclude)

# Compute robust standard errors clustered by school code
robust_se_ols1 <- sqrt(diag(vcovCL(ols_verb1, cluster = ~schlcode)))
robust_se_ols2 <- sqrt(diag(vcovCL(ols_verb2, cluster = ~schlcode)))
robust_se_ols3 <- sqrt(diag(vcovCL(ols_verb3, cluster = ~schlcode)))
robust_se_tsls1 <- sqrt(diag(vcovCL(tsls_verb1, cluster = ~schlcode)))
robust_se_tsls2 <- sqrt(diag(vcovCL(tsls_verb2, cluster = ~schlcode)))

# Generate the table with stargazer
stargazer_verb <- stargazer(
  ols_verb1, ols_verb2, ols_verb3, tsls_verb1, tsls_verb2,
  se = list(robust_se_ols1, robust_se_ols2, robust_se_ols3, robust_se_tsls1, robust_se_tsls2),
  type = 'html',
  title = '<center>Effects of Class Size on Fifth Grade Verbal Scores<center>',
  label = 'tab:class_size_effects_verbal',
  digits = 3,
  align = TRUE,
  dep.var.labels = '<center>Verbal Scores<center>',
  notes = c('Source: Angrist and Lavy (1999)'),
  notes.align = 'l',
  covariate.labels = c('Class size', 'Percent disadvantaged', 'Enrollment', 'Enrollment squared / 100'),
  column.labels = c('OLS', 'OLS', 'OLS', '2SLS', '2SLS'),
  star.cutoffs = c(0.05, 0.01, 0.001),
  omit.stat = c("ser", "adj.rsq")
)
```

```{r, results='asis'}
# Display the results in a well-formatted table
cat('<center>', stargazer_verb, '<center>')
```

