---
title: "Analysis of Union Membership Impact on Wages"
author: "Farnam Taheri" 
date: "`r Sys.Date()`" 
output:
  html_document: # Specifies that the output will be an HTML document
    theme: readable # Uses a theme that is easy to read
    highlight: kate # Syntax highlighting style
    toc: true # Includes a table of contents
    toc_depth: 4 # Depth of headers included in the table of contents
    toc_float: true # Makes the table of contents floating
    df_print: paged # Data frames are printed as paged tables
    code_folding: hide # Does not Allow code blocks to be collapsed
editor_options: 
  markdown: 
    wrap: 72 # Sets the wrap limit for lines in markdown cells
bibliography: HW05.bib # Specifies the bibliography file
---

## [Description]{.smallcaps}

This homework assignment explores the effects of union membership on
workers' wages using a dataset from Vella and Verbeek (1998). We will
use matching techniques to estimate the average treatment effect of
union membership on wages, while controlling for observable
characteristics.

## [Identifying the Effects of a Treatment on an Outcome .]{.smallcaps}

In Identifying the effects of a treatment on an outcome, with CIA
satisfied, we ought to control for the observable characteristics X. We
saw in lecture 9 (regression and causality) how to do so by a parametric
approach: Enter X as independent variables in the regression
specification. Unfortunately, despite many covariates in the
specification, our estimates might still be unreliable. Why? Roughly
speaking, the simple linear regression never differentiates between
observations; it considers the whole untreated subjects as control
group, which is problematic, notably if the data in hand are
observational, where many untreated subjects may not be comparable with
treated ones.

*Matching*, a nonparametric method, helps us overcome this problem. As
noted in the lecture slide, its basic idea is to use observables to
match similar units in different groups and estimate the treatment
effect by averaging the effects between similar units. In this homework
we learn how to implement matching to form a suitable control group and
estimate average treatment effect on the treated in a real world case
discussed during the lecture meetings.

[Vella and Verbeek
1998](https://doi.org/10.1002/(sici)1099-1255(199803/04)13:2%3C163::aid-jae460%3E3.0.co;2-y "Vella, Francis, and Marno Verbeek. 1998. “Whose wages do unions raise? A dynamic model of unionism and wage rate determination for young men.” Journal of Applied Econometrics 13, no. 2 (March): 163–183. ISSN: 1099-1255.")
examine the effects of union membership on workers’ wages, using a panel
data (1980-1987) of 545 young male workers. We want to exploit their
rich data to study the same matter via matching. Please download the
complete
[dataset](https://drive.google.com/file/d/1AF5PqCuF2es1z82uv9PF4QfmASFufohU/view "VV98DATA.zip")
here. This time we use only data of the year 1987, when 143 of workers
are members of unions. You can refer to table 1 of the paper to read the
description of each column of dataset. Below, the steps of carrying out
matching are reviewed and in each step the relevant information for
doing this homework are provided.

### **1.Variables to control.**

The first step in matching is choosing variables that we use to match
treated to untreated subjects. Several factors affect both union
membership status and wage of a worker. In order to make our results
comparable with those presented in page 15 of lecture slide, we consider
the following covariates:

-   1 dummy for rural area

-   3 regional dummies

-   12 sectoral dummies

-   Experience

-   Schooling

### **2. Exact or distance matching.**

Then we determine how close the treatment and control group’s units
should be regarding the selected covariates. The ideal is that they
should be identical - only different in treatment status (union
membership in this problem set). Most probably we cannot find such pairs
in our data, therefore as suggested in the lecture note, we give up
exact matching for the last two covariates: Distance matching for
experience and schooling, exact matching for the others. Exact matching
is often called ”stratification” in the literature.

### **3. How to measure distance.**

Now that we want to do a distance matching for experience and schooling,
we should determine how to measure the closeness of units with respect
to these variables. We use two popular distance measures in this problem
set: Mahalanobis, the propensity score.

### **4. Setting a caliper.**

After specifying the type of distance, we set our criterion for
accepting an untreated subject as a control for a treated one. We should
answer the following question: Does the closest available untreated
subject to a treated one make an acceptable control? If not, we should
set a caliper, i.e., the threshold of closeness of the treatment and
control group. We recognize an untreated unit as a control as long as it
falls within our desired interval based on our selected distance. The
required calipers are specified separately in each question later.

### **5. Pair or multiple matching.**

What if more than one untreated unit were close enough to a treated unit
according to the caliper? Would we regard all of them or a fixed number
of them as controls (multiple matching), or just the closest unit to
that treated one (pair matching)? We employ just pair matching in this
homework.

### **6. Matching with or without replacement.**

Once matched to a treated unit, is the untreated one yet eligible for
becoming matched to other treated units (with replacement) or it is
allowed to be matched to just a single treated unit (without
replacement)? This problem set asks you to do matching without
replacement.

### **7. Selecting a matching method.**

There are several matching methods including nearest 2, optimal, full.
We consider just the first one, which, as mentioned in its help article,
’performs greedy nearest neighbor matching. A distance is computed
between each treated unit and each control unit, and, one by one, each
treated unit is assigned a control unit as a match. The matching is
”greedy” in the sense that there is no action taken to optimize an
overall criterion; Each match is selected without considering the other
matches that may occur subsequently.

### **8. Choosing a matching estimator.**

Up until now a matched sample is constructed. The last step is to choose
a matching estimator. In this homework we simply calculate the average
of differences between outcome (log wage) of treated subjects and that
of their matched subjects, which estimates the treatment effect on the
treated

$$E\left[Y_{1i} - Y_{0i} | D = 1\right] \approx \frac{1}{N_{T_m}} \sum_{i \in T_m} \left[Y_i - Y_{c(i)}\right]$$

This is equal to $δ_1$ in page 14 of the lecture note. In this problem
set, if a treated subject remained without any match, we would ignore
him while computing the above average. Considering the above steps,
answer the following questions:

## [Requirements]{.smallcaps}

### Import required Packages

```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
# Import required packages
library(dplyr) # For data manipulation and transformation
library(tidyr) # For data tidying and reshaping
library(haven) # For reading Stata data (.dta files)
library(tidyverse) # Collection of data manipulation and visualization tools
library(knitr)  # For including R code and results in reports 
library(stargazer)  # For creating publication-quality tables
library(MatchIt)  # For performing matching techniques
library(kableExtra) # For advanced table formatting
library(htmltools) # For HTML formatting and content creation
```

### Set the working directory

```{r message=FALSE, warning=FALSE}
# Set the working directory to the location of the data files
setwd("./Data")

# Assign the working directory to a variable for future reference
working_dir <- getwd()
```

### Load Data

We can load `.RDS` data as follows:

```{r}
# Import the data from the VV98.RDS file as a tibble
VV98_rds <- as_tibble(readRDS(file.path(working_dir, "VV98.RDS"))) |>
  
  # Filter the data to only include observations from the year 1987
  filter(YEAR == 1987) |>
  

  # Select only the necessary columns for the analysis
  select(
    "SCHOOL", "EXPER", "RUR", "NE", "NC",
    "S", "WAGE", "UNION", "AG", "MIN", "CON",
    "TRAD", "TRA", "FIN", "BUS", "PER", "ENT",
    "MAN", "PRO", "PUB"
  )
```

Alternatively, we can import the `.dta` data as follows:

```{r}
# Import the data from the VV98.dta file as a tibble
VV98_dta <- read_stata(file.path(working_dir, "VV98.dta")) |>
  
  # Filter the data to only include observations from the year 1987
  filter(YEAR == 1987) |>
  
  # Select only the necessary columns for the analysis
  select(
    "SCHOOL", "EXPER", "RUR", "NE", "NC",
    "S", "WAGE", "UNION", "AG", "MIN", "CON",
    "TRAD", "TRA", "FIN", "BUS", "PER", "ENT",
    "MAN", "PRO", "PUB"
  )
```

### Renaming variables for clarity

In the following code block we rename the variables in the dataset to
more descriptive names. This aids in understanding the analysis and
ensures that the code is self-explanatory.

```{r}


VV98 <- VV98_rds %>%
  rename(
    YearsOfSchooling = SCHOOL, # Total years of schooling completed
    WorkExperience = EXPER,    # Years of work experience
    RuralArea = RUR,           # Dummy for living in a rural area
    NorthEastRegion = NE,      # Dummy for living in the North East region
    NorthCentralRegion = NC,   # Dummy for living in the North Central region
    SouthernRegion = S,        # Dummy for living in the Southern region
    HourlyWageLog = WAGE,      # Logarithm of the hourly wage
    UnionMember = UNION,       # Dummy for union membership
    AgricultureSector = AG,    # Dummy for working in the agriculture sector
    MiningSector = MIN,        # Dummy for working in the mining sector
    ConstructionSector = CON,  # Dummy for working in the construction sector
    TradeSector = TRAD,        # Dummy for working in the trade sector
    TransportationSector = TRA,# Dummy for working in the transportation sector
    FinanceSector = FIN,       # Dummy for working in the finance sector
    BusinessServiceSector = BUS,# Dummy for working in the business service sector
    PersonalServiceSector = PER,# Dummy for working in the personal service sector
    EntertainmentSector = ENT, # Dummy for working in the entertainment sector
    ManufacturingSector = MAN, # Dummy for working in the manufacturing sector
    ProfessionalServiceSector = PRO,# Dummy for working in the professional service sector
    PublicAdministrationSector = PUB # Dummy for working in the public administration sector
  )
```

## [Questions]{.smallcaps}

### [1. Pre-Matching Analysis: Union Membership and Wage]{.smallcaps}

**Before doing matching, calculate the observed effect of union
membership on wages,** $\mathbb{E}[Y | D = 1] − \mathbb {E}[Y | D = 0]$.

The following code calculates the observed effect of union membership on
wages. It groups the data by union membership status and calculates the
average wage for each group. Then, it computes the difference in average
wages between union and non-union workers.

```{r}
# Calculate the effect of union membership on wages
union_wage_effect <- VV98 %>%
  group_by(UnionMember) %>%  # Group data by union membership status
  summarise(AverageWage = mean(exp(HourlyWageLog))) %>%  # Calculate average wage for each group
  spread(UnionMember, AverageWage) %>%  # Reshape data for comparison
  mutate(ObservedEffect = `1` - `0`)  # Compute the difference in average wages
```

```{r}
# Prepare the results for presentation
wage_comparison_table <- union_wage_effect %>%
  select(`0`, `1`, ObservedEffect) %>%  # Select relevant columns
  rename(                              # Rename for clarity
    'Average Wage (Non-Union)' = `0`,  # Average wage for non-union workers
    'Average Wage (Union)' = `1`,      # Average wage for union workers
    'Union Wage Effect' = ObservedEffect  # Effect of union membership on wages
  )


# Create a formatted table with the results
knitr::kable(wage_comparison_table,
             caption = "<center>Observed Effect of Union Membership on Wages<center>", 
             align = 'c', # Center align the columns
             digits = 3, # Set the number of decimal places to 3
             format = "html", # Output the table in HTML format
             table.attr = "style='width:90%;'", # Set the width of the table
             row.names = FALSE) %>% # Remove row names
        kable_styling(
             bootstrap_options = c("striped", "hover"), # Add Bootstrap styling options
             full_width = F) %>% # Set to false to not stretch the table to full width
        add_footnote( # Add a footnote for the data source
             "Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)") %>%
        scroll_box(width = "100%", height = "500px") # Add a scroll box if the table is too long
```

### [2. Verifying Reliability of the Results: A Covariate Balance Analysis]{.smallcaps}

**Is the previous result reliable? Answer this question via checking out
covariate balance for schooling, experience and the rural dummy; Compare
these variables’ means and standard deviations in the treatment and
control group.**

The following code is used to assess the balance of covariates between
the treatment group (union members) and the control group (non-union
members). Covariate balance is important to ensure that the groups are
comparable on observed characteristics before performing matching.

```{r}
# Assessing Covariate Balance
covariate_balance <- VV98 %>%
  group_by(UnionMember) %>% # Group data by union membership status
  summarise(
    MeanYearsOfSchooling = mean(YearsOfSchooling, na.rm = TRUE),  # Calculate the mean years of schooling, excluding NA values
    SdYearsOfSchooling = sd(YearsOfSchooling, na.rm = TRUE),      # Calculate the standard deviation of years of schooling, excluding NA values
    MeanWorkExperience = mean(WorkExperience, na.rm = TRUE),
    SdWorkExperience = sd(WorkExperience, na.rm = TRUE),
    MeanRuralArea = mean(RuralArea, na.rm = TRUE),
    SdRuralArea = sd(RuralArea, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = -UnionMember, # Exclude the UnionMember column from reshaping
    names_to = "measure", # Create a new column 'measure' for variable names
    values_to = "value"   # Create a new column 'value' for variable values
  ) %>%
  # Reshape the data back to wide format, with separate columns for each UnionMember group
  pivot_wider(
    names_from = UnionMember, # Use the UnionMember values to create new column names
    values_from = value,      # Fill the new columns with values from the 'value' column
    names_glue = "{.value}_{UnionMember}" # Create new column names by combining 'value' and 'UnionMember'
  )

# Display the adjusted covariate balance table
knitr::kable(covariate_balance,
             caption = "<center>Covariate Balance for Schooling, Experience, and Rural Area<center>", 
             align = 'c',
             digits = 3,
             format = "html",
             table.attr = "style='width:90%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F) %>%
  add_footnote("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)")


```

-   **Years of Schooling**:
    -   Mean: The mean years of schooling are slightly higher in the
        control group $(11.786)$ compared to the treatment group
        $(11.713)$. This difference is small and may not be significant.
    -   Standard Deviation (SD): The control group has a higher
        variability in years of schooling $(SD = 1.895)$ compared to the
        treatment group $(SD = 1.248)$.
-   **Work Experience**:
    -   Mean: The mean work experience is almost the same between the
        control group $(10.002)$ and the treatment group (10.049),
        suggesting good balance for this covariate.
    -   SD: The control group shows slightly more variability in work
        experience $(SD = 1.739)$ than the treatment group
        $(SD = 1.396)$.
-   **Rural Area**:
    -   Mean: The proportion of individuals living in rural areas is
        very similar between the control group $(0.221)$ and the
        treatment group $(0.210)$.
    -   SD: The standard deviations are also similar ($0.416$ for the
        control group and 0.409 for the treatment group), indicating
        balanced distribution for this dummy variable.

Overall, the covariate balance for schooling and rural area seems
reasonable, with only minor differences in means and standard
deviations. However, the larger standard deviation for years of
schooling in the control group suggests that there might be more
variability in educational background among non-union members. This
could potentially affect the reliability of the results if not properly
accounted for in the analysis.

### [3. Pair Matching Analysis: No Caliper]{.smallcaps}

**Under pair matching, on a Mahalanobis distance, without a caliper and
without replacement, what is the treatment effect of union membership on
wages of the treated subjects (ATT)?**

This section estimates the Average Treatment Effect on the Treated (ATT)
for union membership on wages using pair matching. The analysis employs
**Mahalanobis** distance without a caliper and without replacement to
find the closest match for each treated subject.

```{r}
# Define the model for pair matching with Mahalanobis distance
pair_matching_model <- matchit(UnionMember ~ YearsOfSchooling + WorkExperience,
                               data = VV98,
                               method = "nearest",
                               distance = "mahalanobis",
                               caliper = NULL,
                               exact = ~ RuralArea + NorthEastRegion + NorthCentralRegion + SouthernRegion + AgricultureSector + MiningSector + ConstructionSector + TradeSector + TransportationSector + FinanceSector + BusinessServiceSector + PersonalServiceSector + EntertainmentSector + ManufacturingSector + ProfessionalServiceSector + PublicAdministrationSector,
                               replace = FALSE)

# Perform the pair matching
matched_pairs <- match.data(pair_matching_model)

# Calculate the ATT by comparing wages between matched pairs
treatment_effect_union_membership <- with(matched_pairs, mean(HourlyWageLog[UnionMember == 1]) - mean(HourlyWageLog[UnionMember == 0]))
```

```{r}
# Create the table with the results
att_result_table <- knitr::kable(data.frame(TreatmentEffectATT = treatment_effect_union_membership),
                                 caption = "<center>Treatment Effect of Union Membership on Wages<center>",
                                 align = 'c',
                                 digits = 3,
                                 col.names = c("Average Treatment Effect on the Treated (ATT)"),
                                 format = "html",
                                 table.attr = "style='width:70%; margin:auto;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                position = "center") %>%
  add_header_above(c(" " = 1)) %>%
  add_footnote(c("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)",
                 "Note: Analysis based on Mahalanobis distance without a caliper and without replacement."))

# Print the ATT result table
att_result_table
```

### [4. Evaluating Matching Performance]{.smallcaps}

**Compare the covariate balance of the initial sample and the matched
sample. According to this, which of the above estimates is preferred?**

In this part, we compare the covariate balance between the initial
sample and the matched sample. Covariate balance ensures that the
treated and control groups are comparable on observed characteristics
before performing matching.

```{r}
# Calculate covariate balance for the initial sample
initial_covariate_balance <- VV98 %>%
  group_by(UnionMember) %>%
  summarise(
    MeanYearsOfSchooling = mean(YearsOfSchooling, na.rm = TRUE),
    SdYearsOfSchooling = sd(YearsOfSchooling, na.rm = TRUE),
    MeanWorkExperience = mean(WorkExperience, na.rm = TRUE),
    SdWorkExperience = sd(WorkExperience, na.rm = TRUE),
    MeanRuralArea = mean(RuralArea, na.rm = TRUE),
    SdRuralArea = sd(RuralArea, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = -UnionMember,
    names_to = "measure",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = UnionMember,
    values_from = value,
    names_glue = "{.value}_{UnionMember}"
  )
```

```{r}
# Calculate covariate balance for the matched sample
matched_covariate_balance <- matched_pairs %>%
  group_by(UnionMember) %>%
  summarise(
    MeanYearsOfSchooling = mean(YearsOfSchooling, na.rm = TRUE),
    SdYearsOfSchooling = sd(YearsOfSchooling, na.rm = TRUE),
    MeanWorkExperience = mean(WorkExperience, na.rm = TRUE),
    SdWorkExperience = sd(WorkExperience, na.rm = TRUE),
    MeanRuralArea = mean(RuralArea, na.rm = TRUE),
    SdRuralArea = sd(RuralArea, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = -UnionMember,
    names_to = "measure",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = UnionMember,
    values_from = value,
    names_glue = "{.value}_{UnionMember}"
  )

```

```{r}
# Table for initial sample covariate balance
initial_covariate_balance_table <- knitr::kable(initial_covariate_balance,
                                                caption = "<center>Covariate Balance for Initial Sample<center>", 
                                                align = 'c',
                                                digits = 3,
                                                format = "html",
                                                table.attr = "style='width:90%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F) %>%
  add_footnote("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)")

```

```{r}
# Table for matched sample covariate balance
matched_covariate_balance_table <- knitr::kable(matched_covariate_balance,
                                                caption = "<center>Covariate Balance for Matched Sample<center>", 
                                                align = 'c',
                                                digits = 3,
                                                format = "html",
                                                table.attr = "style='width:90%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F) %>%
  add_footnote(c("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)",
                 "Note: Analysis based on Mahalanobis distance without a caliper and without replacement."))
```

```{r}
# Print the initial table
initial_covariate_balance_table
```

```{r}
# Print the matcehd table
matched_covariate_balance_table
```

Looking at the means and standard deviations of the covariates in both
samples, we observe that the matched sample has improved balance
compared to the initial sample. Specifically:

-   **Years of Schooling:** The mean and standard deviation of years of
    schooling are closer between the two groups in the matched sample
    compared to the initial sample.

-   **Work Experience:** Similar improvements are seen in the mean and
    standard deviation of work experience.

-   **Rural Area:** The proportions and standard deviations of
    individuals living in rural areas are identical between the treated
    and control groups in both samples.

Overall, the matched sample demonstrates better balance across
covariates, indicating that the matching process has effectively reduced
covariate imbalance between the treated and control groups.

### [5. Support Problem Analysis]{.smallcaps}

**Is our matching exercise in part 3 prone to support problem? Calculate
the percentage of treated subjects for whom a control has been found.
Given this, is our estimate still consistent for ATT?**

In this part, we assess whether the matching exercise is prone to the
support problem. The support problem occurs when there are treated
subjects without suitable controls, which can affect the consistency of
the treatment effect estimate.

```{r}
# Calculate the percentage of treated subjects for whom a control has been found

# Get the number of treated subjects who have been matched
treated_count_0 <- nrow(match.data(pair_matching_model, "treated"))

# Get the total number of subjects who received treatment
total_subjects_treated_count <- nrow(VV98[VV98$UnionMember == 1, ])

# Calculate the percentage of treated subjects who have been matched with a control
percentage_matched_no_caliper <- treated_count_0 / total_subjects_treated_count * 100
```

```{r}
# Create a formatted output for the results
result_message <- paste("Percentage of treated subjects with controls:", round(percentage_matched_no_caliper, 2), "%")

# Check if the estimate for ATT is still consistent based on the percentage of matched treated subjects
consistency_message <- ifelse(percentage_matched_no_caliper < 100, 
                              "Our estimate for ATT might not be fully consistent due to the support problem.",
                              "Our estimate for ATT is likely to be consistent.")

# Create HTML content
html_content <- tags$div(
  style = "text-align: center;", # This will center all the content in the div
  tags$h5(style = "font-weight: bold;", "Support Problem Analysis"), # This will bold the title
  tags$p(result_message),
  tags$p(consistency_message)
)

# Print the HTML content
browsable(html_content)
```

The analysis reveals that there are twice as many controls available for
treated subjects, implying an excess of potential matches. However,
despite this surplus, the estimate for the Average Treatment Effect on
the Treated (ATT) is considered likely to be consistent.

### [6. Pair Matching Analysis: Caliper 1]{.smallcaps}

**Under pair matching, on a Mahalanobis distance, with a caliper of 1
and without replacement, what is the treatment effect of union
membership on wages of the treated subjects (ATT)?**

In this part, we conduct pair matching analysis with a caliper of 1 to
estimate the treatment effect of union membership on wages of the
treated subjects (ATT).

```{r}
# Define the model for pair matching with Mahalanobis distance and a caliper of 1
pair_matching_model_caliper1 <- matchit(UnionMember ~ YearsOfSchooling + WorkExperience,
                                        data = VV98,
                                        method = "nearest",
                                        link = "cauchit",
                                        distance = "mahalanobis",
                                        exact = ~ RuralArea + NorthEastRegion + NorthCentralRegion + SouthernRegion + AgricultureSector + MiningSector + ConstructionSector + TradeSector + TransportationSector + FinanceSector + BusinessServiceSector + PersonalServiceSector + EntertainmentSector + ManufacturingSector + ProfessionalServiceSector + PublicAdministrationSector,
                                        caliper = c(YearsOfSchooling = 1, WorkExperience = 1), # Specify named caliper entries
                                        std.caliper = c(FALSE, FALSE),
                                        replace = FALSE)

# Perform the pair matching with caliper 1
matched_pairs_caliper1 <- match.data(pair_matching_model_caliper1)

# Calculate the ATT by comparing wages between matched pairs with caliper 1
treatment_effect_union_membership_caliper1 <- with(matched_pairs_caliper1, mean(HourlyWageLog[UnionMember == 1]) - mean(HourlyWageLog[UnionMember == 0]))

# Create the table with the results for pair matching with caliper 1
att_result_table_caliper1 <- knitr::kable(data.frame(TreatmentEffectATT = treatment_effect_union_membership_caliper1),
                                          caption = "<center>Treatment Effect of Union Membership on Wages with Caliper 1<center>",
                                          align = 'c',
                                          digits = 3,
                                          col.names = c("Average Treatment Effect on the Treated (ATT)"),
                                          format = "html",
                                          table.attr = "style='width:70%; margin:auto;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                position = "center") %>%
  add_header_above(c(" " = 1)) %>%
  add_footnote(c("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)",
                 "Note: Analysis based on Mahalanobis distance with a caliper of 1 and without replacement."))

# Print the ATT result table for pair matching with caliper 1
att_result_table_caliper1


```

### [7. Optimizing Matching Schemes]{.smallcaps}

**Compare the quality of covariate balance generated by the matching
scheme in part 3 with part 6 . Which one is better? Also compare them
regarding the percentage of treated subjects used in estimation. Explain
the trade-off involved.**

```{r}
# Calculate covariate balance for pair matching with caliper 1
pair_matching_caliper1_covariate_balance <- matched_pairs_caliper1 %>%
  group_by(UnionMember) %>%
  summarise(
    MeanYearsOfSchooling = mean(YearsOfSchooling, na.rm = TRUE),
    SdYearsOfSchooling = sd(YearsOfSchooling, na.rm = TRUE),
    MeanWorkExperience = mean(WorkExperience, na.rm = TRUE),
    SdWorkExperience = sd(WorkExperience, na.rm = TRUE),
    MeanRuralArea = mean(RuralArea, na.rm = TRUE),
    SdRuralArea = sd(RuralArea, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = -UnionMember,
    names_to = "measure",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = UnionMember,
    values_from = value,
    names_glue = "{.value}_{UnionMember}"
  )
```

```{r}
# Table for covariate balance of pair matching without caliper
covariate_balance_table_no_caliper <- knitr::kable(matched_covariate_balance,
                                                  caption = "<center>Covariate Balance for Pair Matching (No Caliper)<center>", 
                                                  align = 'c',
                                                  digits = 3,
                                                  format = "html",
                                                  table.attr = "style='width:90%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F) %>%
  add_footnote("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)")

# Print the covariate balance table for pair matching without caliper
covariate_balance_table_no_caliper


```

```{r}
# Table for covariate balance of pair matching with caliper 1
covariate_balance_table_caliper1 <- knitr::kable(pair_matching_caliper1_covariate_balance,
                                                  caption = "<center>Covariate Balance for Pair Matching (Caliper = 1)<center>", 
                                                  align = 'c',
                                                  digits = 3,
                                                  format = "html",
                                                  table.attr = "style='width:90%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F) %>%
  add_footnote("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)")

# Print the covariate balance table for pair matching with caliper 1
covariate_balance_table_caliper1

```

```{r}
# Calculate the percentage of treated subjects for whom a control has been found

# Get the number of treated subjects who have been matched
treated_count_1 <- nrow(match.data(pair_matching_model_caliper1, "treated"))

# Get the total number of subjects who received treatment
total_subjects_treated_count <- nrow(VV98[VV98$UnionMember == 1, ])

# Calculate the percentage of treated subjects who have been matched with a control
percentage_matched_controls_caliper1 <- treated_count_1 / total_subjects_treated_count * 100

# Create formatted output for the results
result_message_no_caliper <- paste("Percentage of treated subjects used in estimation without a caliper:", round(percentage_matched_no_caliper, 2), "%")
result_message_caliper1 <- paste("Percentage of treated subjects used in estimation with a caliper of 1:", round(percentage_matched_controls_caliper1, 2), "%")

# Create HTML content
html_content <- tags$div(
  style = "text-align: center;", # This will center all the content in the div
  tags$h5(style = "font-weight: bold;", "Percentage of Treated Subjects Used in Estimation"), # This will bold the title
  tags$p(result_message_no_caliper),
  tags$p(result_message_caliper1)
)

# Print the HTML content
browsable(html_content)

```

The results show a comparison of the covariate balance and estimation
efficiency between two pair matching schemes:

1.  Pair matching without a caliper (Part 3)

2.  Pair matching with a caliper of 1 (Part 6)

**Covariate Balance:**

-   Both schemes achieve good covariate balance, with similar means and
    standard deviations for the covariates (Years of Schooling, Work
    Experience, and Rural Area) between the treatment and control
    groups.

-   However, the scheme with a caliper of 1 (Part 6) shows slightly
    better balance, with smaller standard deviations for Years of
    Schooling and Work Experience.

**Estimation Efficiency:**

-   The scheme without a caliper (Part 3) uses 86.01% of the treated
    subjects in estimation, while the scheme with a caliper of 1 (Part
    6)  uses only 65.73% of the treated subjects.
-   This indicates that the scheme with a caliper of 1 is more
    restrictive, discarding more treated subjects from the estimation
    due to the caliper constraint.

**Trade-off:**

-   The trade-off is between covariate balance and estimation
    efficiency.

-   The scheme with a caliper of 1 achieves slightly better covariate
    balance but at the cost of using fewer treated subjects in
    estimation, which may lead to less precise estimates.

-   In contrast, the scheme without a caliper uses more treated
    subjects, potentially leading to more precise estimates, but may
    have slightly worse covariate balance.

### [8. Pair Matching Analysis: Caliper 0.01]{.smallcaps}

**Under pair matching, using the propensity score as distance, with a
caliper of 0.01 and without replacement, calculate ATT.**

In this part, we conduct pair matching analysis with a caliper of 0.01
to estimate the treatment effect of union membership on wages of the
treated subjects (ATT).

```{r}
# Define the model for pair matching with propensity score and a caliper of 0.01
pair_matching_model_caliper_001 <- matchit(UnionMember ~ YearsOfSchooling + WorkExperience,
                                            data = VV98,
                                            method = "nearest",
                                            distance = "glm",
                                            link = "cauchit",
                                            caliper = c(YearsOfSchooling = 0.01, WorkExperience = 0.01), # Specify named caliper entries
                                            std.caliper = c(FALSE, FALSE),
                                            exact = ~ RuralArea + NorthEastRegion + NorthCentralRegion + SouthernRegion + AgricultureSector + MiningSector + ConstructionSector + TradeSector + TransportationSector + FinanceSector + BusinessServiceSector + PersonalServiceSector + EntertainmentSector + ManufacturingSector + ProfessionalServiceSector + PublicAdministrationSector,
                                            replace = FALSE)

# Perform the pair matching with caliper 0.01
matched_pairs_caliper_001 <- match.data(pair_matching_model_caliper_001)

# Calculate the ATT by comparing wages between matched pairs with caliper 0.01
treatment_effect_union_membership_caliper_001 <- with(matched_pairs_caliper_001, mean(HourlyWageLog[UnionMember == 1]) - mean(HourlyWageLog[UnionMember == 0]))

# Create the table with the results for pair matching with caliper 0.01
att_result_table_caliper_001 <- knitr::kable(data.frame(TreatmentEffectATT = treatment_effect_union_membership_caliper_001),
                                              caption = "<center>Treatment Effect of Union Membership on Wages with Caliper 0.01<center>",
                                              align = 'c',
                                              digits = 3,
                                              col.names = c("Average Treatment Effect on the Treated (ATT)"),
                                              format = "html",
                                              table.attr = "style='width:70%; margin:auto;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                position = "center") %>%
  add_header_above(c(" " = 1)) %>%
  add_footnote(c("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)",
                 "Note: Analysis based on propensity score with a caliper of 0.01 and without replacement."))

# Print the ATT result table for pair matching with caliper 0.01
att_result_table_caliper_001
```

### [9. Optimizing Matching Schemes]{.smallcaps}

**Compare the resulting covariate balance with that of part 3. Also
compare them regarding the percentage of treated subjects for whom a
control has been found. According to these comparisons, which one do you
prefer? Argue.**

```{r}
# Calculate covariate balance for pair matching with caliper 0.01 (Part 8)
pair_matching_caliper_001_covariate_balance <- matched_pairs_caliper_001 %>%
  group_by(UnionMember) %>%
  summarise(
    MeanYearsOfSchooling = mean(YearsOfSchooling, na.rm = TRUE),
    SdYearsOfSchooling = sd(YearsOfSchooling, na.rm = TRUE),
    MeanWorkExperience = mean(WorkExperience, na.rm = TRUE),
    SdWorkExperience = sd(WorkExperience, na.rm = TRUE),
    MeanRuralArea = mean(RuralArea, na.rm = TRUE),
    SdRuralArea = sd(RuralArea, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = -UnionMember,
    names_to = "measure",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = UnionMember,
    values_from = value,
    names_glue = "{.value}_{UnionMember}"
  )
# Create tables for covariate balance comparison
covariate_balance_table_part3 <- knitr::kable(initial_covariate_balance,
                                              caption = "<center>Covariate Balance for Pair Matching (No Caliper)<center>", 
                                              align = 'c',
                                              digits = 3,
                                              format = "html",
                                              table.attr = "style='width:90%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F) %>%
  add_footnote("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)")

covariate_balance_table_part8 <- knitr::kable(pair_matching_caliper_001_covariate_balance,
                                              caption = "<center>Covariate Balance for Pair Matching (Caliper = 0.01)<center>", 
                                              align = 'c',
                                              digits = 3,
                                              format = "html",
                                              table.attr = "style='width:90%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F) %>%
  add_footnote("Source: Vella and Verbeek (1998), National Longitudinal Survey (NLS Youth Sample)")
```

```{r}
# Print the tables for covariate balance comparison
covariate_balance_table_part3
```

```{r}
# Print the tables for covariate balance comparison
covariate_balance_table_part8
```

```{r}
# Calculate the percentage of treated subjects for whom a control has been found

# Get the number of treated subjects who have been matched
treated_count_001 <- nrow(match.data(pair_matching_model_caliper_001, "treated"))

# Get the total number of subjects who received treatment
total_subjects_treated_count <- nrow(VV98[VV98$UnionMember == 1, ])

# Calculate the percentage of treated subjects who have been matched with a control
percentage_matched_controls_caliper001 <- treated_count_001 / total_subjects_treated_count * 100

# Create formatted output for the results
result_message_no_caliper <- paste("Percentage of treated subjects used in estimation without a caliper:", round(percentage_matched_no_caliper, 2), "%")
result_message_caliper001 <- paste("Percentage of treated subjects used in estimation with a caliper of 0.01:", round(percentage_matched_controls_caliper001, 2), "%")

# Create HTML content
html_content <- tags$div(
  style = "text-align: center;", # This will center all the content in the div
  tags$h5(style = "font-weight: bold;", "Percentage of Treated Subjects Used in Estimation"), # This will bold the title
  tags$p(result_message_no_caliper),
  tags$p(result_message_caliper001)
)

# Print the HTML content
browsable(html_content)

```

1.  **Covariate Balance:**

    -   Without a caliper, the matching achieves a better balance across
        covariates, as evidenced by the closer means and standard
        deviations between the treated and control groups.

    -   With a caliper of 0.01, while the balance is still acceptable,
        there are cases where the balance is not as tight as without a
        caliper.

2.  **Percentage of Treated Subjects Used in Estimation:**

    -   Without a caliper, a higher percentage (86.01%) of treated
        subjects is utilized in the estimation, providing a more
        comprehensive analysis.

    -   With a caliper of 0.01, the percentage decreases significantly
        to 30.07%, indicating a substantial reduction in the sample
        size, which might lead to less precise estimates.

**Conclusion:**

Considering the trade-offs between covariate balance and sample size,
the matching scheme without a caliper seems preferable.

## [Acknowledgment]{.smallcaps}

This document’s YAML header and output formatting were adapted from
Amirhossein Beykian’s code, with some modifications. A heartfelt thank
you to Amirhossein Beykian for sharing his work, which has significantly
contributed to the clarity, readability, and overall aesthetic quality
of this document. His keen eye for design and willingness to share
knowledge and resources are deeply appreciated, and have helped to
elevate the learning experience for this assignment.

## [References]{.smallcaps}

@Vella1998
